

```{r}

library(Seurat)
library(tidyverse)
library(patchwork)
library(cowplot)
library(hdf5r)
library(gdata)
library(WriteXLS)
library(RColorBrewer)

```

#read data from Joanito et al. 2022 and create Seurat objects
```{r}

PatientData <- readxl::read_xlsx("/dir/Patient_data.xlsx")


EpiMetadata <- readxl::read_xlsx("/dir/EpiMetadata.xlsx")
EpiMetadata <- as.data.frame(EpiMetadata)

EpiMetadata <- EpiMetadata %>% 
  left_join(., PatientData, by=c("patient.ID"))


NonEpiMetadata <- readxl::read_xlsx("/dir/NonEpiMetadata.xlsx")
NonEpiMetadata <- as.data.frame(NonEpiMetadata)

NonEpiMetadata <- NonEpiMetadata %>% 
  left_join(., PatientData, by=c("patient.ID"))


rownames(EpiMetadata) <- (EpiMetadata$cell.ID)
rownames(NonEpiMetadata) <- (NonEpiMetadata$cell.ID)


Epi <- Read10X_h5("/dir/Epithelial_Count_matrix.h5", use.names = TRUE, unique.features = TRUE)
NonEpi <- Read10X_h5("/dir/NonEpithelial_Count_matrix.h5", use.names = TRUE, unique.features = TRUE)

SeuObjEpi <- CreateSeuratObject(counts = Epi, project = "Epithelial",  min.cells = 3, min.features = 200, meta.data = EpiMetadata)
SeuObjNonEpi <- CreateSeuratObject(counts = NonEpi, project = "NonEpithelial",  min.cells = 3, min.features = 200, meta.data = NonEpiMetadata)

SeuObjEpi <- subset(SeuObjEpi, subset = sample.origin == "Normal" )
SeuObjNonEpi <- subset(SeuObjNonEpi, subset = sample.origin == "Normal" )

```

#QC plots on epithelial data
```{r}
 
VlnPlot(SeuObjEpi, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),  ncol = 3, split.by = "orig.ident", pt.size = 0.1)

```

#filter epithelial data
```{r}

SeuObjEpiSub <- subset(SeuObjEpi, subset = nCount_RNA < 25000 & nFeature_RNA < 6000)

VlnPlot(SeuObjEpiSub, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),  ncol = 3, split.by = "orig.ident", pt.size = 0.1)

```

#QC plots on non-epithelial data
```{r}

VlnPlot(SeuObjNonEpi, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),  ncol = 3, split.by = "orig.ident", pt.size = 0.1)

```

#filter non-epithelial data
```{r}

SeuObjNonEpiSub <- subset(SeuObjNonEpi, subset = nCount_RNA < 35000 & nFeature_RNA <4500)

VlnPlot(SeuObjNonEpiSub, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),  ncol = 3, split.by = "orig.ident", pt.size = 0.1)

```

#split combined data into subset based on data origin, normalize, integrate and run PCA
```{r}

CRC16 <- subset(SeuObjNonEpiSub, subset = orig.ident == "CRC16")
CRC16@meta.data[["orig.ident2"]] <- "stroma"

KUL3 <- subset(SeuObjNonEpiSub, subset = orig.ident == "KUL3")
KUL3@meta.data[["orig.ident2"]] <- "stroma"

KUL5 <- subset(SeuObjNonEpiSub, subset = orig.ident == "KUL5")
KUL5@meta.data[["orig.ident2"]] <- "stroma"

SMC <- subset(SeuObjNonEpiSub, subset = orig.ident == "SMC")
SMC@meta.data[["orig.ident2"]] <- "stroma"


Epi_CRC16 <- subset(SeuObjEpiSub, subset = orig.ident == "CRC16")
Epi_CRC16@meta.data[["orig.ident2"]] <- "epi"

Epi_KUL3 <- subset(SeuObjEpiSub, subset = orig.ident == "KUL3")
Epi_KUL3@meta.data[["orig.ident2"]] <- "epi"

Epi_KUL5 <- subset(SeuObjEpiSub, subset = orig.ident == "KUL5")
Epi_KUL5@meta.data[["orig.ident2"]] <- "epi"

Epi_SMC <- subset(SeuObjEpiSub, subset = orig.ident == "SMC")
Epi_SMC@meta.data[["orig.ident2"]] <- "epi"


CRC16 <- merge(CRC16, y = Epi_CRC16)
KUL3 <- merge(KUL3, y = Epi_KUL3)
KUL5 <- merge(KUL5, y = Epi_KUL5)
SMC <- merge(SMC, y = Epi_SMC)


datalist <- list(CRC16, KUL3, KUL5, SMC)


for (i in 1:length(datalist)) {
    datalist[[i]] <- SCTransform(datalist[[i]], vars.to.regress = c("percent.mt"))
    }

features <- SelectIntegrationFeatures(object.list = datalist, nfeatures = 3000)
datalist <- PrepSCTIntegration(object.list = datalist, anchor.features = features)
datalist <- lapply(X = datalist, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = datalist, normalization.method = "SCT",  anchor.features = features, reduction = "rpca", k.anchor = 10)

DataIntegratedHs <- IntegrateData(anchorset = anchors, normalization.method = "SCT")

DefaultAssay(DataIntegratedHs) <- "integrated"

DataIntegratedHs <- RunPCA(DataIntegratedHs, verbose = FALSE)
ElbowPlot(DataIntegratedHs)
DataIntegratedHs <- RunUMAP(DataIntegratedHs, reduction = "pca", dims = 1:25)

DataIntegratedHs <- FindNeighbors(DataIntegratedHs, reduction = "pca", dims = 1:25)
DataIntegratedHs <- FindClusters(DataIntegratedHs, resolution = 0.5)


DefaultAssay(DataIntegratedHs) <- "RNA"

DataIntegratedHs <- NormalizeData(DataIntegratedHs, verbose = F)
DataIntegratedHs <- FindVariableFeatures(DataIntegratedHs, selection.method = "vst", nfeatures = 3000, verbose = FALSE)
DataIntegratedHs <- ScaleData(DataIntegratedHs, verbose = FALSE)

DimPlot(DataIntegratedHs, reduction = "umap", raster =F, label =T)

```

#plot cell type marker genes
```{r}

DefaultAssay(DataIntegratedHs) <- "RNA"

FeaturePlot(DataIntegratedHs, c("PDGFRA", "ACTA2", "RGS5", "VTN", "CSPG4", "PECAM1", "LYVE1", "EPCAM", "S100B", "PTPRC", "FAP"), order = T)

```

#plot data split by epithelial or non-epithelial (stroma) origin
```{r}

DimPlot(DataIntegratedHs, label = F, split.by = "orig.ident2")

```

#plot data split by patient
```{r}

DimPlot(DataIntegratedHs, label = T, split.by = "Patient.ID")

```

#filter out samples from one patient and immune cells continuing with epithelial and mesenchymal cells
```{r}

DataIntegratedHs <- subset(DataIntegrated, invert = T, patient.ID == "CRC2786")

DataIntegrated2Hs <- subset(DataIntegratedHs, idents = c(5,8,10,12,13,14,16,17,24,26))

DefaultAssay(DataIntegrated2Hs) <- "integrated"

DataIntegrated2Hs <- RunPCA(DataIntegrated2Hs, verbose = FALSE)
ElbowPlot(DataIntegrated2Hs)
DataIntegrated2Hs <- RunUMAP(DataIntegrated2Hs, reduction = "pca", dims = 1:20)

DataIntegrated2Hs <- FindNeighbors(DataIntegrated2Hs, reduction = "pca", dims = 1:20)
DataIntegrated2Hs <- FindClusters(DataIntegrated2Hs, resolution = 0.4)

DimPlot(DataIntegrated2Hs, label = T)

```

#plot subsetted data split by epithelial and non-epithelial origin
```{r}

DimPlot(DataIntegrated2Hs, label = F, split.by = "orig.ident2")

```

#find unbiased markers for each subpopulation
```{r}

Markers <- FindAllMarkers(DataIntegrated2Hs, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.25, assay = "RNA")

top10 <- Markers %>%
  top_n(n = 10, wt = avg_log2FC)

DefaultAssay(DataIntegrated2Hs) <- "RNA"

VlnPlot(DataIntegrated2Hs, features = rownames(top10), pt.size = 0)

```

#name populations
```{r}

ClusterIDs <- c("MAF", "SAF","EpithStemProg", "MIF", "Glia", "VEC",
                "SEMF", "Pericyte", "SMC2", "Goblet", "SMC1",
                "LEC", "Colonocyte", "Peri&EC")

names(ClusterIDs) <- levels(DataIntegrated2Hs)
DataIntegrated2Hs <- RenameIdents(DataIntegrated2Hs, ClusterIDs)

DimPlot(DataIntegrated2Hs,
               cols = c("SEMF" = "#F6A937", "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
                        "SMC1" = "#C51B7C", "SMC2" = "#C59ED4", "Pericyte" = "#6E5F57", "Peri&EC" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E", "Glia" = "#C2D9F0",
                        "EpithStemProg" = "#EACAB4",  "Colonocyte" = "#C3936D", "Goblet" = "#734724")) & NoAxes()

```

#filter out presumed doublet cluster
```{r}

DataIntegrated3Hs <- subset(DataIntegrated2Hs, idents = "Peri&EC", invert = T)

DefaultAssay(DataIntegrated3Hs) <- "integrated"

DataIntegrated3Hs <- RunPCA(DataIntegrated3Hs, verbose = FALSE)
DataIntegrated3Hs <- RunUMAP(DataIntegrated3Hs, reduction = "pca", dims = 1:20)

DataIntegrated3Hs <- FindNeighbors(DataIntegrated3Hs, reduction = "pca", dims = 1:20)
DataIntegrated3Hs <- FindClusters(DataIntegrated3Hs, resolution = 0.25)

DimPlot(DataIntegrated3Hs, reduction = "umap", label = T, pt.size = 0.1)

```

#check cell types for clusters
```{r}

DefaultAssay(DataIntegrated3Hs) <- "RNA"

FeaturePlot(DataIntegrated3Hs, c("PDGFRA", "F3", "ADAMDEC1", "PI16", "KCNN3",
                               "RGS5", "PECAM1", "LYVE1", "S100B", "HHIP",
                               "EPCAM", "MUC2"), order = T, ncol = 5) & NoAxes()

```

#rename clusters
```{r}

ClusterIDs <- c( "MAF", 
                 "SAF", "Colonocyte", "MIF", "Glia", "VEC",  "Pericyte", 
                 "SEMF", "SMC", "Goblet", "Myocyte","LEC")

names(ClusterIDs) <- levels(DataIntegrated3Hs)
DataIntegrated4Hs <- RenameIdents(DataIntegrated3Hs, ClusterIDs)

DataIntegrated4Hs@active.ident <- factor(DataIntegrated4Hs@active.ident,
                                       levels = c( "SEMF", "MAF", "SAF",  "MIF",
                                                 "Myocyte", "SMC", "Pericyte", "VEC", "LEC", "Glia",
                                                  "Colonocyte", "Goblet"))

DataIntegrated4Hs$seurat_clusters <- DataIntegrated4Hs@active.ident

saveRDS(DataIntegrated4Hs, file="/dir/Object3_human_colon_integration.rds")

DimPlot(DataIntegrated4Hs,
               cols = c("SEMF" = "#F6A937", "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
                        "Myocyte" = "#C51B7C", "SMC" = "#C59ED4", "Pericyte" = "#47362E", "VEC" = "#E1E0DF", "LEC" = "#A89E99", "Glia" = "#C2D9F0",
                        "EpithStemProg" = "#EACAB4",  "Colonocyte" = "#C3936D", "Goblet" = "#734724")) & NoAxes()

```

#plot data split by dataset of origin
```{r}

DimPlot(DataIntegrated4Hs, split.by = "patient.ID", ncol = 10, group.by = "orig.ident",
        cols = c("SEMF" = "#F6A937", "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
                        "Myocyte" = "#C51B7C", "SMC" = "#C59ED4", "Pericyte" = "#47362E", "VEC" = "#E1E0DF", "LEC" = "#A89E99", "Glia" = "#C2D9F0",
                        "EpithStemProg" = "#EACAB4",  "Colonocytes" = "#C3936D", "Goblet" = "#734724"))

```

#plot fibroblast markers from literature
```{r}

DataIntegrated4Hs@active.ident <- factor(DataIntegrated4Hs@active.ident,
                                       levels = c("Goblet", "Colonocyte","EpithStemProg", "Glia", "LEC",  "VEC", "Pericyte", "SMC","Myocyte", "MIF", "SAF",  "MAF", "SEMF"))

genes <- c("PDGFRA", "ACTA2", "DES", "FOXL1","F3", "PROCR", "BMP5", "BMP7", "WNT5A",  "WNT4", "COL15A1",
           "CD34" , "CD81",  "WNT2B", "GREM1", "GREM2", "RSPO3", "WNT2", "LAMA5", "WIF1","PI16", "KCNN3")

DotPlot(DataIntegrated4Hs, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size")+
    theme(axis.text=element_text(angle=90))

```

#plot markers for non-fibroblast cell types
```{r}

genes <- c("ACTA2",  "LMOD1",  "CNN1", "TAGLN", "MYH11", "MYL9",  "MYLK",  "ACTG2","DES",
            "PDGFRB", "CSPG4" , "NOTCH3",
            "FLT1","PECAM1", 
            "LYVE1",  "FLT4", "PROX1",
            "S100B", "UCHL1",
            "EPCAM", "LGR5", "OLFM4", "CEACAM7", "BEST4", "FABP2", "MUC2")

DotPlot(DataIntegrated4Hs, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size")+
    theme(axis.text=element_text(angle=0))

```

#plot markers for reticular fibroblast and interstitial cells of Cajal
```{r}

FeaturePlot(DataIntegrated4Hs, c("CCL19", "ANO1", "KIT"), order =T) & NoAxes()

```

#plot signatures for stem cell niche factors
```{r}

genelist3 <- list(c("GREM1", "GREM2", "RSPO3"))
DataIntegrated4Hs <- AddModuleScore(object = DataIntegrated4Hs, features = genelist3, name = "genelist3")

FeaturePlot(object = DataIntegrated4Hs, features = "genelist3", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#plot signatures for differentiation niche factors
```{r}

genelist4 <- list(c("BMP5", "BMP7", "WNT5A"))

DataIntegrated4Hs <- AddModuleScore(object = DataIntegrated4Hs, features = genelist4, name = "genelist4")

FeaturePlot(object = DataIntegrated4Hs, features = "genelist4", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#find shared and species-specific markers for human fibroblast subpopulations
```{r}

##SEMF
Markers <- FindMarkers(DataIntegrated4Hs,
                       ident.1 = c("SEMF"),
                       ident.2 = c("Glia", "LEC",  "VEC", "Pericyte", "SMC","Myocyte", "MIF", "SAF",  "MAF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")

top50SEMF <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#go to parallel mouse analysis, get common genes

human_specific <- top50SEMF[!top50SEMF$gene %in% common_genes, ]

top5 <- human_specific %>%
  top_n(n = 5, wt = avg_log2FC)
top5 <- rownames(top5)
HsSEMF <- c(common_genes, top5)


##MAF
Markers <- FindMarkers(DataIntegrated4Hs,
                       ident.1 = c("MAF"),
                       ident.2 = c("Glia", "LEC",  "VEC", "Pericyte", "SMC","yocyte", "MIF", "SAF", "SEMF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")

top50MAF <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#go to parallel mouse analysis, get common genes

human_specific <- top50MAF[!top50MAF$gene %in% common_genes, ]

top5 <- human_specific %>%
  top_n(n = 5, wt = avg_log2FC)

top5 <- rownames(top5)

HsMAF <- c(common_genes, top5)


##SAF
Markers <- FindMarkers(DataIntegrated4Hs,
                       ident.1 = c("SAF"),
                       ident.2 = c("Glia", "LEC",  "VEC", "Pericyte", "SMC","Myocyte", "MIF", "MAF", "SEMF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")

top50SAF <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#go to parallel mouse analysis, get common genes

human_specific <- top50SAF[!top50SAF$gene %in% common_genes, ]

top5 <- human_specific %>%
  top_n(n = 5, wt = avg_log2FC)

top5 <- rownames(top5)

HsSAF <- c(common_genes, top5)


##MIF
Markers <- FindMarkers(DataIntegrated4Hs,
                       ident.1 = c( "MIF"),
                       ident.2 = c("Glia", "LEC",  "VEC", "Pericyte", "SMC","Myocyte", "SAF", "MAF", "SEMF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")


top50MIF <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#go to parallel mouse analysis, get common genes

human_specific <- top50MIF[!top50MIF$gene %in% common_genes, ]

top5 <- human_specific %>%
  top_n(n = 5, wt = avg_log2FC)

top5 <- rownames(top5)

HsMIF <- c(common_genes, top5)



HsPlot <- c(HsSEMF, HsMAF, HsSAF, HsMIF)

DotPlot(DataIntegrated4Hs, assay = "RNA", dot.min = 0.05,  HsPlot, scale.by = "size", col.min = 0)+
    #scale_size(range = c(0.01, 100))+
theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#plot genes that divide mouse fibroblasts into luminal-basal and proximal-distal subtypes
```{r}

DefaultAssay(DataIntegrated4Hs) <- "RNA"

FeaturePlot(DataIntegrated4Hs, c("TNC", "CD55", "WNT4", "LAMA5", "LGR5", "ADAMTS18", "WIF1", "BMP3", "ACTA2"), order =T) & NoAxes()

```

#plot data split by right- and left-sided origin
```{r}

DimPlot(DataIntegrated4Hs, label = F, split.by = "Sidedness")

```

#plot MIF marker genes
```{r}

genes <- c("RAMP1", "KCNN3", "CHODL", "SMOC2", "P2RY1", "THBS2")

DotPlot(DataIntegrated4Hs, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size", col.min = 0)+
    #scale_size(range = c(0.01, 100))+
    theme(axis.text=element_text(angle=90))

```

#make MIF genelist for GSEA
```{r}

MIFsHs <- FindMarkers(DataIntegrated4Hs, ident.1 = "MIF", assay = "RNA", min.pct = 0.15, logfc.threshold = 0.5)
MIFsHs$gene <- rownames(MIFsHs)
MIFsHs <- MIFsHs %>%
  arrange(desc(avg_log2FC))

writexl::write_xlsx(MIFsHs, "/dir/MIFgsea/Hs_MIF_markers_GSEA.xlsx")

```

#GSEA GOMF on MIF
```{r}

library(readr)
library(patchwork)
library(clusterProfiler)
library(org.Mm.eg.db)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(cowplot)
library(readxl)
library(tidyverse)
library(BiocParallel)

Data <- MIFsHs
Data <- Data %>%
  dplyr::select(avg_log2FC, gene)

genelist <- Data$avg_log2FC
names(genelist) <- Data$gene

genelist <- sort(genelist, decreasing = TRUE)

names(genelist)[duplicated(names(genelist))]

head(genelist, 5)
any(genelist == 0)
any(is.na(genelist))
any(names(genelist) == 0)
any(is.na(names(genelist)))

table(is.finite(genelist))

GOMF_set <- msigdbr(species = "Homo sapiens", category = "C5", subcategory = "MF")
GOMFTerm2 <- GOMF_set %>% dplyr::select(gs_name, gene_symbol)

GOMF <- GSEA(
  geneList = genelist,
  TERM2GENE = GOMFTerm2,
  minGSSize = 3,
  pvalueCutoff = 0.05,
  verbose = TRUE,
  BPPARAM = SerialParam())

GOMF@result <- GOMF@result %>%
  arrange(desc(NES))

writexl::write_xlsx(GOMF@result, "/dir/MIFgsea/Hs_GSEA_results.xlsx")

p1 <- gseaplot2(GOMF, geneSetID = "GOMF_GATED_CHANNEL_ACTIVITY", title = "GOMF_GATED_CHANNEL_ACTIVITY")
p2 <- gseaplot2(GOMF, geneSetID = "GOMF_LIGAND_GATED_CHANNEL_ACTIVITY", title = "GOMF_LIGAND_GATED_CHANNEL_ACTIVITY")
p3 <- gseaplot2(GOMF, geneSetID = "GOMF_LIGAND_GATED_MONOATOMIC_CATION_CHANNEL_ACTIVITY", title = "GOMF_LIGAND_GATED_MONOATOMIC_CATION_CHANNEL_ACTIVITY")

gg_p1 <- grid::grid.grabExpr({print(p1)})
gg_p2 <- grid::grid.grabExpr({print(p2)})
gg_p3 <- grid::grid.grabExpr({print(p3)})

wrap_plots(list(gg_p1, gg_p2, gg_p3))

```

```{r}

MIFsHs_ranked <- FindMarkers(DataIntegrated4Hs, ident.1 = "MIF",  logfc.threshold = 0.25, assay = "RNA",)

MIFsHs_ranked$gene <- rownames(MIFsHs_ranked)

MIFsHs_ranked <- MIFsHs_ranked %>%
  arrange(desc(avg_log2FC))

writexl::write_xlsx(MIFsHs_ranked, "/dir/MIFgsea/Hs_MIF_markers_dataset_GSEA.xlsx")

```

```{r}

Data <- MIFsHs_ranked %>%
  dplyr::select(avg_log2FC, gene)
  
genelist <- Data$avg_log2FC
names(genelist) <- Data$gene

genelist <- sort(genelist, decreasing = TRUE)

names(genelist)[duplicated(names(genelist))]

head(genelist, 5)
any(genelist == 0)
any(is.na(genelist))
any(names(genelist) == 0)
any(is.na(names(genelist)))

table(is.finite(genelist))


ICCTerm2 <- read_xlsx("/dir/ICC.xlsx", sheet = 1)
SMCTerm2 <- read_xlsx("/dir/SMC.xlsx", sheet = 1)
MIFTerm2 <- read_xlsx("/dir/MIF.xlsx", sheet = 1)
FIBTerm2 <- read_xlsx("/dir/FIB.xlsx", sheet = 1)
ENSTerm2 <- read_xlsx("/dir/ENS.xlsx", sheet = 1)


ICC <- GSEA(geneList = genelist,
                              TERM2GENE = ICCTerm2,
                              minGSSize = 15,
                              pvalueCutoff = 0.05,
                              verbose = TRUE,
                              scoreType="pos")

MP_SMC <- GSEA(geneList = genelist,
                              TERM2GENE = SMCTerm2,
                              minGSSize = 15,
                              pvalueCutoff = 0.05,
                              verbose = TRUE)

MP_PDGFRa <- GSEA(geneList = genelist,
                              TERM2GENE = MIFTerm2,
                              minGSSize = 15,
                              pvalueCutoff = 0.05,
                              verbose = TRUE,
                              scoreType="pos")

EN <- GSEA(geneList = genelist,
                              TERM2GENE = ENSTerm2,
                              minGSSize = 15,
                              pvalueCutoff = 0.05,
                              verbose = TRUE)

Mucosa_PDGFRa <- GSEA(geneList = genelist,
                              TERM2GENE = FIBTerm2,
                              minGSSize = 15,
                              pvalueCutoff = 0.05,
                              verbose = TRUE,
                              scoreType="pos")

writexl::write_xlsx(Data, "/dir/MIFgsea/Hs_dataset_GSEA_results.xlsx")

NES <- c(MP_PDGFRa@result[["NES"]], Mucosa_PDGFRa@result[["NES"]], ICC@result[["NES"]], MP_SMC@result[["NES"]], EN@result[["NES"]])
CellType <- c("MP_PDGFRa", "Mucosa_PDGFRa", "ICC", "MP_SMC", "EN")

Data <- data.frame(NES, CellType)

Data %>%
  mutate(CellType = fct_relevel(CellType, "MP_PDGFRa", "ICC", "Mucosa_PDGFRa", "MP_SMC", "EN"))%>%
  ggplot(., aes(CellType, NES, fill = CellType))+
  geom_bar(stat = "identity")+
  theme_minimal()+
  scale_fill_manual(values = c("#09A07A", "#0960A0", "#10D8FF", "#697BFB", "#C954F0"))

```

#subset and recluster SMCs
```{r}

DataIntegrated4HsSub <- subset(DataIntegrated4Hs, idents = c("SMC1", "SMC2"))

DefaultAssay(DataIntegrated4HsSub) <- "integrated"

DataIntegrated4HsSub <- RunPCA(DataIntegrated4HsSub, verbose = FALSE)
ElbowPlot(DataIntegrated4HsSub)
DataIntegrated4HsSub <- RunUMAP(DataIntegrated4HsSub, reduction = "pca", dims = 1:10)

DataIntegrated4HsSub <- FindNeighbors(DataIntegrated4HsSub, reduction = "pca", dims = 1:10)
DataIntegrated4HsSub <- FindClusters(DataIntegrated4HsSub, resolution = 0.2)

DimPlot(DataIntegrated4HsSub, reduction = "umap", label = T)

```

#find SMC markers
```{r}

DefaultAssay(DataIntegrated4HsSub) <- "RNA"

Markers <- FindAllMarkers(DataIntegrated4HsSub, only.pos = TRUE, min.pct = 0.4, logfc.threshold = 0.4, min.diff.pct = 0.4, assay = "RNA")

FeaturePlot(DataIntegrated4HsSub, rownames(Markers), order =T) & NoAxes()

```

#rename SMC clusters
```{r}

ClusterIDs <- c( "Myocytes", "MM SMC", "MP SMC", "ICC")
names(ClusterIDs) <- levels(DataIntegrated4HsSub)
DataIntegrated4HsSub <- RenameIdents(DataIntegrated4HsSub, ClusterIDs)

DimPlot(DataIntegrated4HsSub, reduction = "umap", label = F, cols = c("Myocytes" = "#D54B76","MM SMC" = "#F6AEE5", "MP SMC" = "#C59ED4", "ICC" = "#883AA6"), pt.size = 1.4) & NoAxes()

```

#plot SMC markers from literature
```{r}

genes <- c("ACTA2", "MYH11",  "ACTG2", "DES", "HHIP", "KIT")

DefaultAssay(DataIntegrated4HsSub) <- "RNA"

StackedVlnPlot(DataIntegrated4HsSub, genes,  cols = c("Myocytes" = "#D54B76","MM SMC" = "#F6AEE5", "MP SMC" = "#C59ED4", "ICC" = "#883AA6"))

```

#find SMC subtype markers and plot in whole integrated data to find specific ones
```{r}

Markers1 <- FindMarkers(DataIntegrated4HsSub, ident.1 = c("Myocytes"), only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.2, assay = "RNA")
Markers2 <- FindMarkers(DataIntegrated4HsSub, ident.1 = c("MM SMC"), ident.2 = "MP SMC", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.2, assay = "RNA")
Markers3 <- FindMarkers(DataIntegrated4HsSub, ident.1 = c("MP SMC"), ident.2 = "MM SMC", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.2, assay = "RNA")

top1 <- Markers1 %>%
  top_n(n = 50, wt = avg_log2FC)

top2 <- Markers2 %>%
  top_n(n = 50, wt = avg_log2FC)

top3 <- Markers3 %>%
  top_n(n = 50, wt = avg_log2FC)

Genes <- c(rownames(top1), rownames(top2), rownames(top3))


DotPlot(DataIntegrated4Hs, assay = "RNA", dot.min = 0.05, Genes, scale.by = "size", col.min = 0)+
    #scale_size(range = c(0.01, 100))+
theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#plot SMC subtype markers in subsetted data
```{r, fig.height=2.5, fig.width=10, dpi =200}

DotPlot(DataIntegrated4HsSub, assay = "RNA", dot.min = 0.05,c("HHIP", "NPNT", "SOSTDC1", "DIO2", "GREM2", "RSPO2", "CHRDL1", "SULF1","SPOCK1"), scale.by = "size", col.min = 0)+
    theme(axis.text.x=element_text(angle=90))

```

#Spearman's correlation between mouse and human fubroblast subtypes
```{r}

library(biomaRt)
library(pheatmap)
library(AnnotationDbi)
library(gprofiler2)
library(homologene)
library(ComplexHeatmap)
library(circlize)
library(nichenetr)

```

```{r}

DataIntegrated4 <- readRDS("/dir/Object2_mouse_colon_integration.rds")

obj_mouse <- subset(DataIntegrated4, idents = c("Luminal_SEMF", "Basal_SEMF", "Luminal_MAF",  "Prox_basal_MAF", "Dist_basal_MAF", "SMAF", "MIF"))
ClusterIDs <- c( "Mm_Luminal_SEMF", "Mm_Basal_SEMF", "Mm_Luminal_MAF",  "Mm_Prox_basal_MAF", "Mm_Dist_basal_MAF", "Mm_SAF", "Mm_MIF")
names(ClusterIDs) <- levels(obj_mouse)
obj_mouse <- RenameIdents(obj_mouse, ClusterIDs)
obj_mouse$seurat_clusters <- obj_mouse@active.ident
DefaultAssay(obj_mouse) <- "RNA"

obj_human <- subset(DataIntegrated4Hs, idents = c("SEMF", "MAF", "SMAF",  "MIF"))
ClusterIDs <- c("Hs_SEMF", "Hs_MAF", "Hs_SAF",  "Hs_MIF")
names(ClusterIDs) <- levels(obj_human)
obj_human <- RenameIdents(obj_human, ClusterIDs)
obj_human$seurat_clusters <- obj_human@active.ident
DefaultAssay(obj_human) <- "RNA"

obj_human2 <- FindVariableFeatures(obj_human, nfeatures = 1000)
obj_mouse2 <- FindVariableFeatures(obj_mouse, nfeatures = 1000)

human_var <- VariableFeatures(obj_human2)

mouse_symbols <- convert_human_to_mouse_symbols(human_var)

human_to_mouse <- data.frame(
  human = human_var,
  mouse = mouse_symbols,
  stringsAsFactors = FALSE
)

human_to_mouse <- human_to_mouse[
  !is.na(human_to_mouse$mouse),
]

human_to_mouse <- human_to_mouse[
  !duplicated(human_to_mouse$mouse),
]

human_mat <- GetAssayData(
  obj_human2,
  assay = "RNA",
  slot = "data"
)

mouse_mat <- GetAssayData(
  obj_mouse2,
  assay = "RNA",
  slot = "data"
)

human_mat <- human_mat[
  human_to_mouse$human,
]

rownames(human_mat) <- human_to_mouse$mouse

shared_genes <- intersect(
  rownames(mouse_mat),
  rownames(human_mat)
)

cat("Number of shared genes:", length(shared_genes), "\n")

human_mat <- human_mat[shared_genes, ]
mouse_mat <- mouse_mat[shared_genes, ]

human_avg <- AverageExpression(
  obj_human2,
  assays = "RNA",
  slot = "data",
  features = human_to_mouse$human,
  group.by = "seurat_clusters"
)$RNA

human_avg <- human_avg[
  human_to_mouse$human,
]

rownames(human_avg) <- human_to_mouse$mouse

human_avg <- human_avg[
  shared_genes,
]

mouse_avg <- AverageExpression(
  obj_mouse2,
  assays = "RNA",
  slot = "data",
  features = shared_genes,
  group.by = "seurat_clusters"
)$RNA

human_avg <- as.matrix(human_avg)
mouse_avg <- as.matrix(mouse_avg)

cor_matrix <- cor(mouse_avg,
  human_avg,
  method = "spearman")

```

```{r, fig.height=4, fig.width=5}

pheatmap(cor_matrix, clustering_method = "ward.D2", display_numbers = TRUE, border_color = NA)

```

```{r, fig.height=4, fig.width=5}

breaks <- c(
  seq(0.50, 0.62, length.out = 67), 
  seq(0.62, 0.75, length.out = 33)  )

pheatmap(cor_matrix, clustering_method = "ward.D2", display_numbers = TRUE, border_color = NA, breaks = breaks)

```

