---
title: "Code4_human_CRC_normal_integration_scRNAseq"
output: html_document
date: "2025-07-14"
editor_options: 
  chunk_output_type: inline
---
#load libraries
```{r}

library(tidyverse)
library(patchwork)
library(cowplot)
library(hdf5r)
library(gdata)
library(remotes)
library(Seurat)
library(uwot)
library(ggpubr)
library(WriteXLS)
library(RColorBrewer)
library(monocle3)
library(monocle)

```

#read data from Joanito et al. 2022 and create Seurat objects
```{r}

PatientData <- readxl::read_xlsx("/dir/Patient_data.xlsx")


EpiMetadata <- readxl::read_xlsx("/dir/EpiMetadata.xlsx")
EpiMetadata <- as.data.frame(EpiMetadata)


EpiMetadata <- EpiMetadata %>% 
  left_join(., PatientData, by=c("patient.ID"))


NonEpiMetadata <- readxl::read_xlsx("/dir/NonEpiMetadata.xlsx")
NonEpiMetadata <- as.data.frame(NonEpiMetadata)


NonEpiMetadata <- NonEpiMetadata %>% 
  left_join(., PatientData, by=c("patient.ID"))

rownames(EpiMetadata) <- (EpiMetadata$cell.ID)

rownames(NonEpiMetadata) <- (NonEpiMetadata$cell.ID)

Epi <- Read10X_h5("/dir/Epithelial_Count_matrix.h5", use.names = TRUE, unique.features = TRUE)
NonEpi <- Read10X_h5("/dir/NonEpithelial_Count_matrix.h5", use.names = TRUE, unique.features = TRUE)


SeuObjEpi <- CreateSeuratObject(counts = Epi, project = "Epithelial",  min.cells = 3, min.features = 200, meta.data = EpiMetadata)
SeuObjNonEpi <- CreateSeuratObject(counts = NonEpi, project = "NonEpithelial",  min.cells = 3, min.features = 200, meta.data = NonEpiMetadata)

SeuObjEpi <- subset(SeuObjEpi, subset = sample.origin == "Normal" | sample.origin == "Tumor" | sample.origin == "Tumor-2")
SeuObjNonEpi <- subset(SeuObjNonEpi, subset = sample.origin == "Normal" | sample.origin == "Tumor" | sample.origin == "Tumor-2")

SeuObjEpi@meta.data[["sample.origin"]][SeuObjEpi@meta.data[["sample.origin"]] == "Tumor-2"] <- "Tumor"
SeuObjNonEpi@meta.data[["sample.origin"]][SeuObjNonEpi@meta.data[["sample.origin"]] == "Tumor-2"] <- "Tumor"

```

#QC plots on epithelial data
```{r}
 
VlnPlot(SeuObjEpi, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),  ncol=3)

```

#filter epithelial data
```{r}

SeuObjEpiSub <- subset(SeuObjEpi, subset = nCount_RNA < 50000)

VlnPlot(SeuObjEpiSub, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),  ncol=3)

```

#QC plots on non-epithelial data
```{r}

VlnPlot(SeuObjNonEpi, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),  ncol=3)

```

#filter non-epithelial data
```{r}

SeuObjNonEpiSub <- subset(SeuObjNonEpi, subset = nCount_RNA < 70000 & nFeature_RNA <5500)

VlnPlot(SeuObjNonEpiSub, features = c("nCount_RNA", "nFeature_RNA"),  ncol=2)

```

#split combined data into subset based on data origin, normalize, integrate and run PCA
```{r}

CRC16 <- subset(SeuObjNonEpiSub, subset = orig.ident == "CRC16")
CRC16@meta.data[["orig.ident2"]] <- "stroma"

KUL3 <- subset(SeuObjNonEpiSub, subset = orig.ident == "KUL3")
KUL3@meta.data[["orig.ident2"]] <- "stroma"

KUL5 <- subset(SeuObjNonEpiSub, subset = orig.ident == "KUL5")
KUL5@meta.data[["orig.ident2"]] <- "stroma"

SMC <- subset(SeuObjNonEpiSub, subset = orig.ident == "SMC")
SMC@meta.data[["orig.ident2"]] <- "stroma"

JSC <- subset(SeuObjNonEpiSub, subset = orig.ident == "JSC")
JSC@meta.data[["orig.ident2"]] <- "stroma"


Epi_CRC16 <- subset(SeuObjEpiSub, subset = orig.ident == "CRC16")
Epi_CRC16@meta.data[["orig.ident2"]] <- "epi"

Epi_KUL3 <- subset(SeuObjEpiSub, subset = orig.ident == "KUL3")
Epi_KUL3@meta.data[["orig.ident2"]] <- "epi"

Epi_KUL5 <- subset(SeuObjEpiSub, subset = orig.ident == "KUL5")
Epi_KUL5@meta.data[["orig.ident2"]] <- "epi"

Epi_SMC <- subset(SeuObjEpiSub, subset = orig.ident == "SMC")
Epi_SMC@meta.data[["orig.ident2"]] <- "epi"

Epi_JSC <- subset(SeuObjEpiSub, subset = orig.ident == "JSC")
Epi_JSC@meta.data[["orig.ident2"]] <- "epi"


CRC16 <- merge(CRC16, y = Epi_CRC16)
KUL3 <- merge(KUL3, y = Epi_KUL3)
KUL5 <- merge(KUL5, y = Epi_KUL5)
SMC <- merge(SMC, y = Epi_SMC)
JSC <- merge(JSC, y = Epi_JSC)

datalist <- list(CRC16, KUL3, KUL5, SMC, JSC)

for (i in 1:length(datalist)) {
    datalist[[i]] <- SCTransform(datalist[[i]], vars.to.regress = c("percent.mt"))
    }

features <- SelectIntegrationFeatures(object.list = datalist, nfeatures = 3000)
datalist <- PrepSCTIntegration(object.list = datalist, anchor.features = features)
datalist <- lapply(X = datalist, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = datalist, normalization.method = "SCT",  anchor.features = features, reduction = "rpca", k.anchor = 10)
DataIntegrated <- IntegrateData(anchorset = anchors, normalization.method = "SCT")

DataIntegrated <- subset(DataIntegrated, patient.ID == "CRC2786", invert =T)

DefaultAssay(DataIntegrated) <- "integrated"

DataIntegrated <- RunPCA(DataIntegrated, verbose = FALSE)
ElbowPlot(DataIntegrated)
DataIntegrated <- RunUMAP(DataIntegrated, reduction = "pca", dims = 1:30)

DataIntegrated <- FindNeighbors(DataIntegrated, reduction = "pca", dims = 1:30)
DataIntegrated <- FindClusters(DataIntegrated, resolution = 0.57)


DefaultAssay(DataIntegrated) <- "RNA"

DataIntegrated <- NormalizeData(DataIntegrated, verbose = F)
DataIntegrated <- FindVariableFeatures(DataIntegrated, selection.method = "vst", nfeatures = 3000, verbose = FALSE)
DataIntegrated <- ScaleData(DataIntegrated, verbose = FALSE)

DimPlot(DataIntegrated, reduction = "umap", raster =F, label =T)

```

#plot cell type marker genes
```{r}

genes <- c("PDGFRA", "FAP", "DES", 
           "ACTA2", "DES", "CSPG4", "RGS5", "PECAM1", "LYVE1", "S100B",
           "EPCAM",  "MUC2", "CA1","LGR5",
           "PTPRC", "C1QA", "CXCL8", "FCER1A", "KIT",
           "MKI67", "IL7R", "CD3D", "CD4",  "CD8A", "MKI67",  "FOXP3",
            "CD79A", "CD19", "IGHA1", "IGHG1", "IGKC")

FeaturePlot(DataIntegrated, genes, order =T)

```

#find unbiased markers for each subpopulation
```{r}

Markers <- FindAllMarkers(DataIntegrated, assay = "RNA")

genes <- Markers %>%
  top_n(n = 10, wt = avg_log2FC)

VlnPlot(DataIntegrated, rownames(genes), pt.size = 0)

```

#name populations
```{r}

ClusterIDs <- c("CD8", "CD4",    "tissueMacr","CD8", "iCMS2Epi",  "B", 
                "Treg", "CAF", "iCMS3Epi", "IgA_Plasma_IGKC+", "Endoth",
                "CD4", "IgA_Plasma_IGKC+", "Fib", "IgG_Plasma_IGKC+", "Macr",
                "Fib", "Goblet", "Peri",   "IgA_Plasma",   "IgA_Plasma", 
                "Colonocyte","DC", "IgG_Plasma", "glia", "Mast", 
                "Fib","Neutr","SMC", "proT", "KI67+immune")
names(ClusterIDs) <- levels(DataIntegrated)
DataIntegrated2 <- RenameIdents(DataIntegrated, ClusterIDs)


DimPlot(DataIntegrated2, label = F, raster =F,
         cols = c( "Fib" = "#05B985",  "CAF" = "#FA32D2", 'SMC' = '#C59ED4',
             "Peri" =  "#F6F00D" ,
             "Endoth" = "#69AF0D",
             "glia" = "#F37907",
             "Goblet" = "#A55910", "Colonocyte" = "#61350B",  "iCMS2Epi" = "#CC8745", "iCMS3Epi" = "#F7C08A",
             "Macr" = "#131313",  "tissueMacr" = "#5D5B5A", "DC" = "#92908D","Neutr" = "#BEB9B5", "Mast" = "#E1DDDA",
             "proT" = "#DEECEF","CD4" = "#9ED2DD", "CD8" = "#4DB6CC",   "Treg" = "#1E90A7",  "KI67+immune" = "#045565",
             "B" = "#5A7B99",
             "IgA_Plasma" = "#E3DBAE", "IgG_Plasma" = "#F9EB96","IgA_Plasma_IGKC+" = "#EBD56F", "IgG_Plasma_IGKC+" = "#EFC502")) & NoAxes()

```

#subset stroma (mesenchymal cells) for further annotation
```{r}

Stroma <- subset(DataIntegrated2, idents = c("Fib", "SMC", "CAF", "Peri", "Endoth", "glia"))

DefaultAssay(Stroma) <- "integrated"

Stroma <- RunPCA(Stroma, verbose = FALSE)
ElbowPlot(Stroma)
Stroma <- RunUMAP(Stroma, reduction = "pca", dims = 1:15)

Stroma <- FindNeighbors(Stroma, reduction = "pca", dims = 1:15)
Stroma <- FindClusters(Stroma, resolution = 0.5)

DimPlot(Stroma, label = TRUE)

```

#plot stromal cell type marker genes
```{r}

DefaultAssay(Stroma) <- "RNA"

FeaturePlot(Stroma, c("PDGFRA", "KCNN3", "HHIP", "DES", "NOTCH3", "PECAM1", "LYVE1", "S100B", "KIT", "CCL19", "CXCL14",
                      "VSTM2A", "SFTA1P", "PI16", "KCNN3", 
                      "MMP1", "CXCL1", "IL6",
                      "MMP11", "COL10A1", "COL11A1",
                      "FAP", "MYH11", "ACTA2", "LRRC15",
                       #artery ECs
                      "GJA5", "FBLN5", 
                      #capillary ECs 
                      "CA4", "CD36",  "BTNL9",
                      #vein ECs
                      "ACKR1", "SELP"), order = T)

```

#find unbiased markers for stromal subpopulations
```{r}

Markers <- FindAllMarkers(Stroma, assay = "RNA")

genes <- Markers %>%
  top_n(n = 10, wt = avg_log2FC)

VlnPlot(Stroma, rownames(genes), pt.size = 0)

```

#name stromal populations
```{r}

ClusterIDs <- c("mCAF", "MAF", "angPeri", "SAF", "preCAF2", "glia",
                 "MIF",  "venEC", "tipEC", "art_capEC","contrPeri",
                "preCAF1", "SEMF", "SMC","iCAF", "lymphEC",
                "myocytes")

names(ClusterIDs) <- levels(Stroma)
Stroma <- RenameIdents(Stroma, ClusterIDs)

DimPlot(Stroma, label = TRUE)

```

#use stromal subcluster names to annotate same populations in the whole data
```{r}

Stroma@meta.data$populations <- Stroma@active.ident
FibLabels <- select(Stroma@meta.data, c(cell.ID, populations))

DataIntegrated2CR <- DataIntegrated2
DataIntegrated2CRC@meta.data[["annotation"]] <- DataIntegrated2CRC@active.ident


CellLabels <- select(DataIntegrated2CRC@meta.data, c(cell.ID, annotation))

Labels <- left_join(CellLabels, FibLabels, by = "cell.ID" )


fill_char_values <- function(df, col_to_fill, reference_col) {
  df[[col_to_fill]] <- ifelse(is.na(df[[col_to_fill]]),
                              as.character(df[[reference_col]]),  # Convert to character if needed
                             as.character(df[[col_to_fill]]))
  return(df)
}

Labels <- fill_char_values(Labels, "populations", "annotation")
Labels <- Labels[,c(1,3)]

DataIntegrated2CRC@meta.data <- left_join(DataIntegrated2CRC@meta.data, Labels, by = "cell.ID" )
DataIntegrated2CRC@meta.data$populations.y <- as.factor(DataIntegrated2CRC@meta.data$populations.y)

Idents(DataIntegrated2CRC) <- "populations.y"

row.names(DataIntegrated2CRC@meta.data) <- DataIntegrated2CRC@meta.data$cell.ID


DataIntegrated2CRC@active.ident <- factor(DataIntegrated2CRC@active.ident,
                                       levels = c("SEMF","MAF","SAF", "MIF", "myocytes", "SMC",
                                                 "preCAF1", "preCAF2", "iCAF",  "mCAF",
                                                 "contrPeri", "angPeri", "lymphEC", "venEC", "art_capEC", "tipEC", "glia",
                                                  "Goblet", "Colonocyte",  "iCMS2Epi", "iCMS3Epi",
                                                   "Macr",  "tissueMacr", "DC","Neutr", "Mast", 
                                                  "proT","CD4", "CD8",   "Treg",  "KI67+immune",
                                                  "B",  "IgA_Plasma", "IgG_Plasma","IgA_Plasma_IGKC+", "IgG_Plasma_IGKC+"))


DimPlot(DataIntegrated2CRC, label = F, raster =F,
         cols = c("SEMF" = "#F6A937",  "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
                  "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
                  "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
                  "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
                  "lymphEC" = "#1C7522",
                  "venEC" = "#099E46",  "art_capEC" = "#1DE00B",  "tipEC" = "#AFF233",
                  "glia" = "#F37907",
                  "Goblet" = "#A55910", "Colonocyte" = "#61350B",  "iCMS2Epi" = "#CC8745", "iCMS3Epi" = "#F7C08A",
                  "Macr" = "#131313",  "tissueMacr" = "#5D5B5A", "DC" = "#92908D","Neutr" = "#BEB9B5", "Mast" = "#E1DDDA",
                  "proT" = "#DEECEF","CD4" = "#9ED2DD", "CD8" = "#4DB6CC",   "Treg" = "#1E90A7",  "KI67+immune" = "#045565",
                  "B" = "#5A7B99",
                  "IgA_Plasma" = "#E3DBAE", "IgG_Plasma" = "#F9EB96","IgA_Plasma_IGKC+" = "#EBD56F", "IgG_Plasma_IGKC+" = "#B79702")) & NoAxes()

```

#plot data split by dataset of origin
```{r}

DimPlot(DataIntegrated2CRC, label = F, raster =F,
         cols = c("SEMF" = "#F6A937",  "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
                  "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
                  "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
                  "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
                  "lymphEC" = "#1C7522",
                  "venEC" = "#099E46",  "art_capEC" = "#1DE00B",  "tipEC" = "#AFF233",
                  "glia" = "#F37907",
                  "Goblet" = "#A55910", "Colonocyte" = "#61350B",  "iCMS2Epi" = "#CC8745", "iCMS3Epi" = "#F7C08A",
                  "Macr" = "#131313",  "tissueMacr" = "#5D5B5A", "DC" = "#92908D","Neutr" = "#BEB9B5", "Mast" = "#E1DDDA",
                  "proT" = "#DEECEF","CD4" = "#9ED2DD", "CD8" = "#4DB6CC",   "Treg" = "#1E90A7",  "KI67+immune" = "#045565",
                  "B" = "#5A7B99",
                  "IgA_Plasma" = "#E3DBAE", "IgG_Plasma" = "#F9EB96","IgA_Plasma_IGKC+" = "#EBD56F", "IgG_Plasma_IGKC+" = "#B79702"),
        split.by = "orig.ident") & NoAxes()


```

#plot non-mesenchymal cell type markers
```{r}

DataIntegrated2CRC@active.ident <- factor(DataIntegrated2CRC@active.ident,
                                          levels = c("IgG_Plasma_IGKC+", "IgA_Plasma_IGKC+",
                                                     "IgG_Plasma","IgA_Plasma",
                                                     "B", "Treg", "CD8", "CD4", "proT", "KI67+immune", 
                                                     "Mast", "DC", "Neutr", "tissueMacr", "Macr", 
                                                     "iCMS3Epi", "iCMS2Epi", "Colonocyte","Goblet",
                                                     "glia",  
                                                     "tipEC", "art_capEC", "venEC","lymphEC", 
                                                     "angPeri", "contrPeri", "SMC", "myocytes",
                                                     "mCAF", "iCAF", "preCAF2", "preCAF1", 
                                                     "MIF", "SAF","MAF", "SEMF"))

genes <- c("EPCAM",  "MUC2", "CA1","LGR5", "AXIN2", "CTSE", "MMP7",
           "PTPRC", "IL1B",
           "C1QA", "CXCL8", "FCER1A",
           "MKI67", "IL7R", "CD3D", "CD8A", "CD4",  "FOXP3",
            "CD79A", "CD19", "IGHA1", "IGHG1", "IGKC")

DotPlot(DataIntegrated2CRC,  assay = "RNA", dot.min = 0.05,  genes, scale.by = "size", col.min = 0)+
      theme(axis.text=element_text(angle=90))

```

#plot data split by sample type (normal/tumor)
```{r}

DimPlot(DataIntegrated2CRC, label = F, raster =F,
         cols = c("SEMF" = "#F6A937",  "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
                  "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
                  "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
                  "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
                  "lymphEC" = "#1C7522",
                  "venEC" = "#099E46",  "art_capEC" = "#1DE00B",  "tipEC" = "#AFF233",
                  "glia" = "#F37907",
                  "Goblet" = "#A55910", "Colonocyte" = "#61350B",  "iCMS2Epi" = "#CC8745", "iCMS3Epi" = "#F7C08A",
                  "Macr" = "#131313",  "tissueMacr" = "#5D5B5A", "DC" = "#92908D","Neutr" = "#BEB9B5", "Mast" = "#E1DDDA",
                  "proT" = "#DEECEF","CD4" = "#9ED2DD", "CD8" = "#4DB6CC",   "Treg" = "#1E90A7",  "KI67+immune" = "#045565",
                  "B" = "#5A7B99",
                  "IgA_Plasma" = "#E3DBAE", "IgG_Plasma" = "#F9EB96","IgA_Plasma_IGKC+" = "#EBD56F", "IgG_Plasma_IGKC+" = "#B79702"),
        split.by = "sample.origin") & NoAxes()


```

#continue with stromal (mesenchymal) cells for downstream analysis
```{r}

DefaultAssay(Stroma) <- "integrated"

ElbowPlot(Stroma)

Stroma <- RunPCA(Stroma, verbose = FALSE)
Stroma <- RunUMAP(Stroma, reduction = "pca", dims = 1:15)

Stroma <- FindNeighbors(Stroma, reduction = "pca", dims = 1:15)
Stroma <- FindClusters(Stroma, resolution = 0.6)

DimPlot(Stroma, label = TRUE)

DefaultAssay(Stroma) <- "RNA"

FeaturePlot(Stroma, c("PECAM1", "RGS5", order = T))
#Filter out endoth + rgs5 doublet cluster
Stroma <- subset(Stroma, idents = 18, invert = T)

DefaultAssay(Stroma) <- "integrated"

ElbowPlot(Stroma)

Stroma <- RunPCA(Stroma, verbose = FALSE)
Stroma <- RunUMAP(Stroma, reduction = "pca", dims = 1:15)

Stroma <- FindNeighbors(Stroma, reduction = "pca", dims = 1:15)
Stroma <- FindClusters(Stroma, resolution = 0.4)

DimPlot(Stroma, label = TRUE)

```

#filter out samples form stroma subanalysis: leave out patients who contribute to less than 5 clusters
```{r}

clusters_per_sample <- table(Stroma@meta.data$patient.ID, Stroma@meta.data$populations)

# Count the number of clusters for each sample
clusters_per_sample <- rowSums(clusters_per_sample > 0)

# Identify samples that contribute cells to a minimum of 5 clusters
samples_to_keep <- names(clusters_per_sample[clusters_per_sample >= 5])

# Subset the Seurat object to include only samples that contribute to a minimum of 5 clusters
Stroma <- subset(Stroma, subset = patient.ID %in% samples_to_keep)

DimPlot(Stroma, split.by = "patient.ID", ncol = 6)

```

#filter out patients who contribute less than 50 cells in total to the stroma analysis
```{r}

sample_names <-Stroma@meta.data[["patient.ID"]]

# Calculate the number of cells per sample
cell_counts <- tapply(1:nrow(Stroma@meta.data), Stroma@meta.data$patient.ID, length)

# Identify samples with less than 50 cells
samples_to_remove <- sample_names[cell_counts[sample_names] < 50]
samples_to_remove <- unique(samples_to_remove)

Idents(Stroma) <- "patient.ID"

Stroma <- subset(Stroma, idents = samples_to_remove, invert = T)

Idents(Stroma) <- "seurat_clusters"

DimPlot(Stroma, label = T)

```

#re-cluster after filtering out cells
```{r}

DefaultAssay(Stroma) <- "integrated"

ElbowPlot(Stroma)
Stroma <- RunPCA(Stroma, verbose = FALSE)
Stroma <- RunUMAP(Stroma, reduction = "pca", dims = 1:26)
Stroma <- FindNeighbors(Stroma, reduction = "pca", dims = 1:26)
Stroma <- FindClusters(Stroma, resolution = 0.6)

DimPlot(Stroma, label = TRUE)

```

#identify and name stromal cell subtypes
```{r}

DotPlot(Stroma,  assay = "RNA", dot.min = 0.05,
        c("PDGFRA", "VSTM2A",   "SFTA1P",  "PI16", "KCNN3",
          "FAP", "WNT5A","CXCL14", "HGF",
          "PDGFRL","C3", "SFRP2", "CTGF", "COL1A1",
           "MMP1" , "CXCL1", "IL6", 
           "MMP11", "HOPX", "COL10A1", "COL11A1",
          "HHIP", "DES",
           "PDGFRB", "NOTCH3", "CSPG4", "RERGL", "RGS5",
          "ACTA2", "CNN1", "TAGLN", "MYH11",  "MYL9", "MYLK", "ACTG2", "TPM1", "TPM2",  "LUM",
          "LYVE1","PECAM1",  "SELP", "CD36", "CA4",  "APLN", "ESM1", "S100B"),
        scale.by = "size", col.min = 0)+
  theme(axis.text=element_text(angle=90))


ClusterIDs <- c("MAF",  "preCAF1",  "angPeri",  "SAF", "preCAF2", "tipEC",
                "mCAF",  "glia", "MIF", "venEC", "art_capEC",
               "contrPeri", "SEMF", "iCAF", "SMC",  "lymphEC",
                  "myocytes")
names(ClusterIDs) <- levels(Stroma)
Stroma <- RenameIdents(Stroma, ClusterIDs)


Stroma@active.ident <- factor(Stroma@active.ident,
                                       levels = c("SEMF", "MAF", "SAF", "MIF",
                                                 "preCAF1", "preCAF2", "iCAF",  "mCAF",
                                                  "myocytes", "SMC", 
                                                 "contrPeri","angPeri",
                                                  "art_capEC", "venEC", "tipEC",
                                                  "lymphEC", "glia"))

DimPlot(Stroma, label = TRUE) & NoAxes()

DimPlot(Stroma,  cols = c("SEMF" = "#F6A937",  "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
            "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
                            "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
                         "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
                         "lymphEC" = "#1C7522",
                         "venEC" = "#099E46",  "art_capEC" = "#1DE00B",  "tipEC" = "#AFF233",
                         "glia" = "#F37907")) & NoAxes()


```

#plot data split by dataset of origin
```{r}

DimPlot(Stroma,  cols = c(  "SEMF" = "#F6A937",  "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
         
            "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
                            "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
                         "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
                         "lymphEC" = "#1C7522",
                         "venEC" = "#099E46",  "art_capEC" = "#1DE00B",  "tipEC" = "#AFF233",
                         "glia" = "#F37907"), split.by = "orig.ident") & NoAxes()

```

#plot data split by patient
```{r}

DimPlot(Stroma,  cols = c("SEMF" = "#F6A937",  "MAF" = "#1D7BFC", 'SAF' = '#129B93',  "MIF" = "#035C56",
         
            "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
                            "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
                         "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
                         "lymphEC" = "#1C7522",
                         "venEC" = "#099E46",  "art_capEC" = "#1DE00B",  "tipEC" = "#AFF233",
                         "glia" = "#F37907"), split.by = "patient.ID") & NoAxes()

```

#plot proportions of stromal cells by patient
```{r}

Patients <- select(Stroma@meta.data, c(patient.ID, populations, sample.origin))

Patients <- Patients %>%
  group_by(patient.ID, populations, sample.origin) %>% 
  summarise(Nb = n()) %>%
  ungroup()%>%
  group_by(patient.ID, sample.origin) %>% 
  mutate(Count = sum(Nb)) %>%
  mutate(Percent = Nb/Count*100) %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  ungroup()

PatientData <- unique(select(Stroma@meta.data, c(patient.ID, sample.origin)))

Patients <- left_join(Patients, PatientData, by = c("patient.ID", "sample.origin"))

Patients %>%
  mutate(category_reordered = factor(populations, levels = c(
    "SEMF", "MAF", "SAF", "MIF", 
    "preCAF1", "preCAF2", "iCAF", "mCAF",
    "myocytes", "SMC",
    "contrPeri", "angPeri",
    "lymphEC", "venEC", "art_capEC", "tipEC",
    "glia"
  ))) %>% 
  ggplot(aes(x = patient.ID, y = Percent, fill = category_reordered)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_grid( ~ sample.origin, scales = "free", space = "free")+
  scale_fill_manual(values = c(
    "SEMF" = "#F6A937", "MAF" = "#1D7BFC", "SAF" = "#129B93", "MIF" = "#035C56",
    "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
    "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
    "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
    "lymphEC" = "#1C7522", "venEC" = "#099E46", "art_capEC" = "#1DE00B", "tipEC" = "#AFF233",
    "glia" = "#F37907"
  )) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#plot markers for interstitial ceells of Cajal, reticular fibroblasts, proliferative CAFs and apCAFs
```{r}

FeaturePlot(Stroma, c("KIT", "CCL19", "MKI67", "CD74"), order = T, ncol = 4) & NoAxes()

```

#plot mesenchymal cell type markers
```{r}

Stroma@active.ident <- factor(Stroma@active.ident,levels = c( "glia", "tipEC", "art_capEC","venEC",  "lymphEC", 
                                                               "angPeri", "contrPeri", "SMC",  "myocytes",
                                                              "mCAF",  "iCAF",  "preCAF2","preCAF1",
                                                               "MIF","SAF",  "MAF", "SEMF"))
 
DotPlot(Stroma,  assay = "RNA", dot.min = 0.05,
        c("PDGFRA", "VSTM2A",   "SFTA1P",  "PI16", "KCNN3",
          "FAP", "WNT5A","CXCL14", "HGF",
          "PDGFRL","C3", "SFRP2", "CTGF", "COL1A1",
           "MMP1" , "CXCL1", "IL6", 
           "MMP11", "HOPX", "COL10A1", "COL11A1",
          "HHIP", "DES",
           "PDGFRB", "NOTCH3", "CSPG4", "RERGL", "RGS5",
          "ACTA2", "CNN1", "TAGLN", "MYH11",  "MYL9", "MYLK", "ACTG2", "TPM1", "TPM2",  "LUM",
          "LYVE1","PECAM1",  "SELP", "CD36", "CA4",  "APLN", "ESM1", "S100B"),
        scale.by = "size", col.min = 0)+
  theme(axis.text=element_text(angle=90))

```

#plot mesenchymal cell proportions in normal tissue-derived samples
```{r}

StromaStage <- select(Stroma@meta.data, c(patient.ID, populations, sample.origin, Group.Stage))

StromaStage <- StromaStage %>%
  filter(sample.origin == "Normal") %>%
  group_by(sample.origin, populations) %>%
  summarise(Nb = n()) %>%
  mutate(Total = sum(Nb)) %>%
  ungroup() %>%
  mutate(Percent = Nb / Total * 100) %>%
  mutate(across(where(is.numeric), round, 2))

StromaStage %>%
  mutate(category_reordered = factor(populations, levels = c(
  "myocytes", "SMC",
    "contrPeri", "angPeri",
    "lymphEC", "venEC", "art_capEC", "tipEC",
    "glia",
    "SEMF", "MAF", "SAF", "MIF", 
    "preCAF1", "preCAF2", "iCAF", "mCAF"
  ))) %>%
  ggplot(aes(x = sample.origin, y = Percent, fill = category_reordered)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c(
    "SEMF" = "#F6A937", "MAF" = "#1D7BFC", "SAF" = "#129B93", "MIF" = "#035C56",
    "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
    "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
    "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
    "lymphEC" = "#1C7522", "venEC" = "#099E46", "art_capEC" = "#1DE00B", "tipEC" = "#AFF233",
    "glia" = "#F37907"
  )) +
  theme_minimal() +
  xlab("Tumor Stage Group") +
  ylab("Percent Composition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#plot mesenchymal cell proportions in tumor tissue-derived samples split by early and late stage
```{r}

StromaNormal <- select(Stroma@meta.data, c(patient.ID, populations, sample.origin, Group.Stage))

StromaNormal <- StromaNormal %>%
  filter(sample.origin == "Tumor") %>%
  mutate(StageGroup = ifelse(Group.Stage %in% c("1", "2"), "1/2", "3/4")) %>%
  group_by(StageGroup, populations) %>%
  summarise(Nb = n()) %>%
  group_by(StageGroup) %>%
  mutate(Total = sum(Nb)) %>%
  ungroup() %>%
  mutate(Percent = Nb / Total * 100) %>%
  mutate(across(where(is.numeric), round, 2))

StromaNormal %>%
  mutate(category_reordered = factor(populations, levels = c(
  "myocytes", "SMC",
    "contrPeri", "angPeri",
    "lymphEC", "venEC", "art_capEC", "tipEC",
    "glia",
    "SEMF", "MAF", "SAF", "MIF", 
    "preCAF1", "preCAF2", "iCAF", "mCAF"
  ))) %>%
  ggplot(aes(x = StageGroup, y = Percent, fill = category_reordered)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c(
    "SEMF" = "#F6A937", "MAF" = "#1D7BFC", "SAF" = "#129B93", "MIF" = "#035C56",
    "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#7012E5",
    "myocytes" = "#C51B7C", "SMC" = "#C59ED4",
    "contrPeri" = "#F6F00D", "angPeri" = "#E2C513",
    "lymphEC" = "#1C7522", "venEC" = "#099E46", "art_capEC" = "#1DE00B", "tipEC" = "#AFF233",
    "glia" = "#F37907"
  )) +
  theme_minimal() +
  xlab("Tumor Stage Group") +
  ylab("Percent Composition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#find markerd specific to all CAFs
```{r}

panCAF <- FindMarkers(DataIntegrated2, ident.1 = c("preCAF1", "preCAF2", "iCAF", "mCAF"), ident.2 = c("SEMF", "MAF", "SAF", "MIF"), assay = "RNA", logfc.threshold = 0.25, min.pct = 0.25, min.diff.pct = 0.45, only.pos = T)

DotPlot(DataIntegrated2CRC, assay = "RNA", dot.min = 0.05,  
        c("FAP", "CTHRC1", "INHBA",  "THBS2", "WNT2", "PODNL1","CPXM1", "ADAM12",  "COL11A1","SPHK1", "BMP1", 
          "SULF2", "RAB31","COL8A1", "CHPF", "PLOD2", "KIAA1217", "PRRX1", "RAI14", "RUNX1", "LMCD1", "TNFRSF12A", "POSTN",  "SOX4",  "BGN", "CYTOR", "MFAP2",  "CDH11", "RCN3"),
        scale.by = "size")+
    theme(axis.text.x=element_text(angle=90))

```

#find marekrs for each CAF subtype
```{r}

preCAF1 <- FindMarkers(Stroma, ident.1 = "preCAF1", assay = "RNA", logfc.threshold = 0.25, min.pct = 0.25, min.diff.pct = 0.4, only.pos = T)
preCAF2 <- FindMarkers(Stroma, ident.1 = "preCAF2", assay = "RNA", logfc.threshold = 0.25, min.pct = 0.25, min.diff.pct = 0.4, only.pos = T)
iCAF <- FindMarkers(Stroma, ident.1 = "iCAF", assay = "RNA", logfc.threshold = 0.25, min.pct = 0.25, min.diff.pct = 0.4, only.pos = T)
mCAF <- FindMarkers(Stroma, ident.1 = "mCAF", assay = "RNA", logfc.threshold = 0.25, min.pct = 0.25, min.diff.pct = 0.4, only.pos = T)

```

#plot preCAF1 markers
```{r}

preCAF1 <- filter(preCAF1, Name %in%  c( "F2RL2", "CPXM1",   "MME", "TMEM158", "COL7A1",  "CXCL1","FAM20C", "PDPN","WNT2", 
                                         "CST1","DIO2", "TNFSF11",
                                         "TRPA1", "SOX6","HSD17B2", "NSG1","TSPAN33",  "AGT", "F3",
                                         "DLL1", "EMID1", "SMPDL3A", "CXCL14", "PTCH1","PTGER2", "CYGB"))

DotPlot(DataIntegrated2CRC,  assay = "RNA", dot.min = 0.05,  rownames(preCAF1), scale.by = "size", col.min = 0)+
      theme(axis.text=element_text(angle=90))

```

#plot preCAF2 markers
```{r}

preCAF2 <- filter(preCAF2, Name %in% c("IGF1", "GAS1", "CYP1B1","RARRES1","CRABP2",   "COMP", "SFRP4",  "COL8A1",  "ADAM12",
             "MYH10", "FNDC1", "ZFHX4",  "STEAP1", "STEAP2" , "GFPT2", "PDGFRL",  "BOC",
             "C3", "CCDC80", "PRELP","OGN","PODN", "PTGIS", "PTGDS", "MMP23B",
             "SFRP2",  "CILP",  "SLIT3",   "MEDAG" ))

DotPlot(DataIntegrated2CRC,  assay = "RNA", dot.min = 0.05,  rownames(preCAF2), scale.by = "size", col.min = 0)+
      theme(axis.text=element_text(angle=90))

```

#plot iCAF markers
```{r}

iCAF <- filter(iCAF, Name %in%  c("MMP1", "MMP3", "CXCL5", "CXCL6",
            "CA12", "CXCL1", "CXCL3", "GK",  "CXCL8","TNFAIP6", "IL7R",
            "SPHK1", "PDPN",  "CHI3L1", "HSD11B1",  "FAM20C",
            "PTGES", "TMEM132A", "TMEM158", "COL7A1", "WNT5A", "MME",
            "GFPT2", "STEAP1", "STEAP2" ))

DotPlot(DataIntegrated2CRC,  assay = "RNA", dot.min = 0.05,  rownames(iCAF), scale.by = "size", col.min = 0)+
      theme(axis.text=element_text(angle=90))

```

#plot mCAF markers
```{r}

mCAF <- filter(mCAF, Name %in% c("COL10A1", "COL11A1","PLPP4", "GJB2",   "NOX4", "NTM",  "BHLHE41", "HOPX",
             "LINC01705",
             "COMP", "ITGBL1",  "SLC12A8", 
               "SFRP4",   "COL16A1", "ITGA11","MMP11", "ASPN"))

DotPlot(DataIntegrated2CRC,  assay = "RNA", dot.min = 0.05,  rownames(mCAF), scale.by = "size", col.min = 0)+
      theme(axis.text=element_text(angle=90))

```

#plot pan-CAF and subset specific CAF markers
```{r}

DotPlot(DataIntegrated2CRC, c("ACTA2", "PDGFRA", "PDGFRB", "PDPN", "POSTN", "LRRC15", "TNC","S100A4",
                           "FAP", 
                           "CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                           "F2RL2","TSPAN33", "TMEM158",
                           "IGF1", "GAS1","CRABP2",   "COMP", 
                           "MMP1", "MMP3", "CXCL5", "CXCL6",
                           "COL10A1", "COL11A1","PLPP4"), assay = "RNA", dot.min = 0.05,  scale.by = "size", col.min = 0)+
  theme(axis.text.x=element_text(angle=90))

```

#plot pan-CAF marker expression in TCGA-COAD data
```{r}

library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(ggplot2)
library(ggpubr)

query <- GDCquery(
  project = "TCGA-COAD",
  data.category = "Transcriptome Profiling",
  data.type = "Gene Expression Quantification",
  workflow.type = "STAR - Counts",
  sample.type = c("Primary Tumor", "Solid Tissue Normal")
)

GDCdownload(query)
se <- GDCprepare(query)  
se

rowdat <- rowData(se)

symbol_col <- dplyr::case_when(
  "external_gene_name" %in% colnames(rowdat) ~ "external_gene_name",
  "gene_name" %in% colnames(rowdat) ~ "gene_name",
  "symbol" %in% colnames(rowdat) ~ "symbol",
  TRUE ~ NA_character_
)
stopifnot(!is.na(symbol_col))

signature_genes <- c("FAP", "CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1")

idx <- which(rowdat[[symbol_col]] %in% signature_genes)
stopifnot(length(idx) > 0)

libsize <- colSums(counts)
cpm_mat <- sweep(counts[idx, , drop=FALSE], 2, libsize, "/") * 1e6
log_expr <- log2(cpm_mat + 1)

plot_df <- as.data.frame(t(log_expr))
plot_df$sample <- rownames(plot_df)

gene_symbols <- rowdat[[symbol_col]][idx]
colnames(plot_df)[1:length(idx)] <- make.unique(gene_symbols)

plot_df <- plot_df %>%
  left_join(
    data.frame(
      sample = colnames(counts),
      group = ifelse(coldat$sample_type == "Solid Tissue Normal", "Normal adjacent", "CRC tumor")
    ),
    by = "sample"
  ) %>%
  pivot_longer(
    cols = -c(sample, group),
    names_to = "gene",
    values_to = "expr"
  )

plot_df$group <- factor(plot_df$group, levels = c("Normal adjacent", "CRC tumor"), labels = c("Normal", "Tumor"))
plot_df$gene  <- factor(plot_df$gene, levels = signature_genes)

ggplot(plot_df, aes(group, expr)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.35, size = 0.7) +
  stat_compare_means(method = "t.test", label = "p.format") +
  facet_wrap(~ gene, ncol = 7) +   # <-- same y-axis
  labs(
    title = "panCAF genes (TCGA-COAD)",
    y = "log2(CPM + 1)",
    x = ""
  ) +
  theme_classic()

```

#plot signature plot of pan-CAF markers
```{r}

genelist5 <- list(c("FAP", "CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1"))
Stroma <- AddModuleScore(object = Stroma, features = genelist5, name = "genelist5")

FeaturePlot(object = Stroma, features = "genelist5", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#correlation analysis of FAP and pan-CAF marker expression in TCGA-COAD data
```{r}

library(TCGAbiolinks)
library(SummarizedExperiment)
library(edgeR)      
library(ggplot2)
library(ggpubr)      
library(dplyr)
library(tidyr)
library(biomaRt)

query <- GDCquery(
  project = "TCGA-COAD",
  data.category = "Transcriptome Profiling",
  data.type = "Gene Expression Quantification",
  workflow.type = "STAR - Counts"
)

GDCdownload(query)
coad_data <- GDCprepare(query)


counts <- assay(coad_data)

dge <- DGEList(counts = counts)
dge <- calcNormFactors(dge)
log2_cpm <- cpm(dge, log = TRUE, prior.count = 1)


rownames(log2_cpm) <- sub("\\..*$", "", rownames(log2_cpm))


genes_of_interest <- c("FAP", "CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1")
ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")

gene_map <- getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol"),
  filters = "hgnc_symbol",
  values = genes_of_interest,
  mart = ensembl)


genes_present <- gene_map$ensembl_gene_id
expr_subset <- log2_cpm[rownames(log2_cpm) %in% genes_present, , drop = FALSE]

expr_subset_t <- t(expr_subset)
colnames(expr_subset_t) <- gene_map$hgnc_symbol[match(colnames(expr_subset_t), gene_map$ensembl_gene_id)]
expr_df <- as.data.frame(expr_subset_t)


target_gene <- "FAP"
other_genes <- c("CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1")

other_genes_present <- intersect(other_genes, colnames(expr_df))

expr_long <- expr_df %>%
  dplyr::select(all_of(c(target_gene, other_genes_present))) %>%
  pivot_longer(
    cols = all_of(other_genes_present),
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  mutate(GeneLabel = paste0(Gene))

expr_long$GeneLabel <- factor(expr_long$GeneLabel,
                              levels = c("CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1"))

ggplot(expr_long, aes(x = !!sym(target_gene), y = Expression)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(method = "lm", color = "red", se = FALSE, size = 0.75) +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")), method = "pearson", size = 3) +
  facet_wrap(~GeneLabel, scales = "free_y", ncol = 6) +
  theme_minimal(base_size = 15) +
  labs(x = paste0(target_gene), y = "mRNA expression (log2)")

```

#expression of FAP/CTHRC1 in different CMSs of CRC in Sidra-LUMC AC-ICAM NatMed 2023 data
```{r}

#FAP mRNA expression data was downloaded from Sidra-LUMC AC-ICAM NatMed 2023 from cBioPortal

df <- read_table("FAP.txt")

comparisons <- list(
  c("CMS1", "CMS4"),
  c("CMS2", "CMS4"),
  c("CMS3", "CMS4"),
  c("CMS4", "mixed"))

ggplot(df, aes(x = CMS, y = expression, fill = CMS)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  stat_compare_means(comparisons = comparisons, 
                     method = "t.test", 
                     label = "p.format", 
                     p.adjust.method = "bonferroni") +
  theme_minimal() +
  labs(title = "FAP",
        y = "mRNA expression normalized counts") +
  theme(legend.position = "none")+
  scale_fill_manual(values = c("#eff3ff","#bdd7e7","#6baed6","#2171b5","#aaaaaa"))

```

```{r}

#CTHRC1 mRNA expression data was downloaded from Sidra-LUMC AC-ICAM NatMed 2023 from cBioPortal

df <- read_table("CTHRC1.txt")

comparisons <- list(
  c("CMS1", "CMS4"),
  c("CMS2", "CMS4"),
  c("CMS3", "CMS4"),
  c("CMS4", "mixed"))

ggplot(df, aes(x = CMS, y = expression, fill = CMS)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  stat_compare_means(comparisons = comparisons, 
                     method = "t.test", 
                     label = "p.format", 
                     p.adjust.method = "bonferroni") +
  theme_minimal() +
  labs(title = "FAP",
        y = "mRNA expression normalized counts") +
  theme(legend.position = "none")+
  scale_fill_manual(values = c("#eff3ff","#bdd7e7","#6baed6","#2171b5","#aaaaaa"))

```

#correlation analysis of FAP/CTHRC1 and iCAF/mCAF marker expression in TCGA-COAD data
```{r}

genes_of_interest <- c("FAP", "CTHRC1", "COL10A1", "COL11A1", "PLPP4", "MMP1", "CXCL5", "CXCL6")
ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")

gene_map <- getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol"),
  filters = "hgnc_symbol",
  values = genes_of_interest,
  mart = ensembl)

genes_present <- gene_map$ensembl_gene_id
expr_subset <- log2_cpm[rownames(log2_cpm) %in% genes_present, , drop = FALSE]

expr_subset_t <- t(expr_subset)
colnames(expr_subset_t) <- gene_map$hgnc_symbol[match(colnames(expr_subset_t), gene_map$ensembl_gene_id)]
expr_df <- as.data.frame(expr_subset_t)

target_gene <- "FAP"
other_genes <- c("COL10A1", "COL11A1", "PLPP4", "MMP1", "CXCL5", "CXCL6")

other_genes_present <- intersect(other_genes, colnames(expr_df))

expr_long <- expr_df %>%
  dplyr::select(all_of(c(target_gene, other_genes_present))) %>%
  pivot_longer(
    cols = all_of(other_genes_present),
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  mutate(GeneLabel = paste0(Gene))

expr_long$GeneLabel <- factor(expr_long$GeneLabel,
                              levels = c("COL10A1", "COL11A1", "PLPP4", "MMP1", "CXCL5", "CXCL6"))

ggplot(expr_long, aes(x = !!sym(target_gene), y = Expression)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(method = "lm", color = "red", se = FALSE, size = 0.75) +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")), method = "pearson", size = 3) +
  facet_wrap(~GeneLabel, scales = "free_y", ncol = 6) +
  theme_minimal(base_size = 15) +
  labs(x = paste0(target_gene), y = "mRNA expression (log2)")

```

```{r}

target_gene <- "CTHRC1"
other_genes <- c("COL10A1", "COL11A1", "PLPP4", "MMP1", "CXCL5", "CXCL6")

other_genes_present <- intersect(other_genes, colnames(expr_df))

expr_long <- expr_df %>%
  dplyr::select(all_of(c(target_gene, other_genes_present))) %>%
  pivot_longer(
    cols = all_of(other_genes_present),
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  mutate(GeneLabel = paste0(Gene))

expr_long$GeneLabel <- factor(expr_long$GeneLabel,
                              levels = c("COL10A1", "COL11A1", "PLPP4", "MMP1", "CXCL5", "CXCL6"))

ggplot(expr_long, aes(x = !!sym(target_gene), y = Expression)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(method = "lm", color = "red", se = FALSE, size = 0.75) +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")), method = "pearson", size = 3) +
  facet_wrap(~GeneLabel, scales = "free_y", ncol = 6) +
  theme_minimal(base_size = 15) +
  labs(x = paste0(target_gene), y = "mRNA expression (log2)")

```

#Monocle3 trajectory analysis on mesenchymal cells
```{r}

cds <- SeuratWrappers::as.cell_data_set(Stroma)
cds <- cluster_cells(cds, resolution=0.00012)

p1 <- plot_cells(cds, color_cells_by = "cluster", show_trajectory_graph = FALSE) + NoAxes()
p2 <- plot_cells(cds, color_cells_by = "partition", show_trajectory_graph = FALSE) + NoAxes()

wrap_plots(p1, p2) 

```

```{r}

cds <- learn_graph(cds, use_partition = TRUE, verbose = FALSE)

plot_cells(cds,
           color_cells_by = "cluster",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE) + NoAxes()

```

```{r}
cds <- order_cells(cds, root_cells = colnames(cds[,clusters(cds) == c(6,8, 14)]))

plot_cells(cds,
           color_cells_by = "pseudotime",
           group_cells_by = "cluster",
           label_cell_groups = FALSE,
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           label_roots = FALSE,
           trajectory_graph_color = "grey60") + NoAxes()

```

#signature plot of expression of candidate transcription factor identified in CellOracle analysis
```{r}

genelist6 <- list(c("PRRX1", "MAFB", "TWIST1"))
Stroma <- AddModuleScore(object = Stroma, features = genelist6, name = "genelist6")

FeaturePlot(object = Stroma, features = "genelist6", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#correlation analysis on candidate TF and CAF marker gene expression
```{r}

genes_of_interest <- c("PRRX1",
                       "FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                       "MMP1", "CXCL5", "CXCL6", "COL10A1", "COL11A1", "PLPP4")

ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")

gene_map <- getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol"),
  filters = "hgnc_symbol",
  values = genes_of_interest,
  mart = ensembl)

genes_present <- gene_map$ensembl_gene_id
expr_subset <- log2_cpm[rownames(log2_cpm) %in% genes_present, , drop = FALSE]

expr_subset_t <- t(expr_subset)
colnames(expr_subset_t) <- gene_map$hgnc_symbol[match(colnames(expr_subset_t), gene_map$ensembl_gene_id)]
expr_df <- as.data.frame(expr_subset_t)


target_gene <- "PRRX1"
other_genes <- c( "FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                  "MMP1", "CXCL5", "CXCL6", "COL10A1", "COL11A1", "PLPP4")

other_genes_present <- intersect(other_genes, colnames(expr_df))

expr_long <- expr_df %>%
  dplyr::select(all_of(c(target_gene, other_genes_present))) %>%
  pivot_longer(
    cols = all_of(other_genes_present),
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  mutate(GeneLabel = paste0(Gene))

cor_df <- expr_long %>%
  group_by(GeneLabel) %>%
  summarise(
    R = cor(Expression, .data[[target_gene]], use = "pairwise.complete.obs"),
    p = cor.test(Expression, .data[[target_gene]], method = "pearson")$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p, method = "BH"))

gene_order <- c("FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                  "MMP1", "CXCL5", "CXCL6", "COL10A1", "COL11A1", "PLPP4")

cor_df <- cor_df %>%
  mutate(
    GeneLabel = factor(GeneLabel, levels = gene_order),
    sig_label = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ "ns"
    )
  )

ggplot(cor_df, aes(x = GeneLabel, y = R)) +
  geom_col(width = 0.8) +
  geom_text(
    aes(label = sig_label),
    vjust = ifelse(cor_df$R >= 0, -0.3, 1.3),
    size = 5
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  theme_minimal(base_size = 15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "PRRX1", y = "Pearson R")

```

```{r}

target_gene <- "TWIST1"
other_genes <- c( "FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                  "MMP1", "CXCL5", "CXCL6","COL10A1", "COL11A1", "PLPP4")

other_genes_present <- intersect(other_genes, colnames(expr_df))

expr_long <- expr_df %>%
  dplyr::select(all_of(c(target_gene, other_genes_present))) %>%
  pivot_longer(
    cols = all_of(other_genes_present),
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  mutate(GeneLabel = paste0(Gene))

cor_df <- expr_long %>%
  group_by(GeneLabel) %>%
  summarise(
    R = cor(Expression, .data[[target_gene]], use = "pairwise.complete.obs"),
    p = cor.test(Expression, .data[[target_gene]], method = "pearson")$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p, method = "BH"))   

gene_order <- c( "FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                  "MMP1", "CXCL5", "CXCL6", "COL10A1", "COL11A1", "PLPP4")

cor_df <- cor_df %>%
  mutate(
    GeneLabel = factor(GeneLabel, levels = gene_order),
    sig_label = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ "ns"
    )
  )

ggplot(cor_df, aes(x = GeneLabel, y = R)) +
  geom_col(width = 0.8) +
  geom_text(
    aes(label = sig_label),
    vjust = ifelse(cor_df$R >= 0, -0.3, 1.3),
    size = 5
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  theme_minimal(base_size = 15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "TWIST1", y = "Pearson R")

```

```{r}

target_gene <- "MAFB"
other_genes <- c( "FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                  "MMP1", "CXCL5", "CXCL6", "COL10A1", "COL11A1", "PLPP4")

other_genes_present <- intersect(other_genes, colnames(expr_df))

expr_long <- expr_df %>%
  dplyr::select(all_of(c(target_gene, other_genes_present))) %>%
  pivot_longer(
    cols = all_of(other_genes_present),
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  mutate(GeneLabel = paste0(Gene))

cor_df <- expr_long %>%
  group_by(GeneLabel) %>%
  summarise(
    R = cor(Expression, .data[[target_gene]], use = "pairwise.complete.obs"),
    p = cor.test(Expression, .data[[target_gene]], method = "pearson")$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p, method = "BH"))   # optional FDR

gene_order <- c("FAP","CTHRC1", "INHBA", "THBS2", "WNT2", "PODNL1", "CPXM1",
                  "MMP1", "CXCL5", "CXCL6", "COL10A1", "COL11A1", "PLPP4")

cor_df <- cor_df %>%
  mutate(
    GeneLabel = factor(GeneLabel, levels = gene_order),
    sig_label = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ "ns"
    )
  )

ggplot(cor_df, aes(x = GeneLabel, y = R)) +
  geom_col(width = 0.8) +
  geom_text(
    aes(label = sig_label),
    vjust = ifelse(cor_df$R >= 0, -0.3, 1.3),
    size = 5
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4) +
  theme_minimal(base_size = 15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "MAFB", y = "Pearson R")

```

#subsetting CAFs for abundance analysis
```{r}

Stroma@meta.data$populations <- Stroma@active.ident

Tumor <- subset(Stroma, sample.origin == "Tumor")

CAFs <- subset(Tumor, idents = c("preCAF1", "preCAF2", "mCAF", "iCAF"))

CAFs2 <- CAFs
CAFs2@meta.data$CAFpopulations <- CAFs2@active.ident
sample_names <-CAFs2@meta.data[["patient.ID"]]

cell_counts <- tapply(1:nrow(CAFs2@meta.data), CAFs2@meta.data$patient.ID, length)

samples_to_remove <- sample_names[cell_counts[sample_names] < 50]
samples_to_remove <- unique(samples_to_remove)

Idents(CAFs2) <- "patient.ID"

CAFs2 <- subset(CAFs2, idents = samples_to_remove, invert = T)

Idents(CAFs2) <- "CAFpopulations"


PatientPlots <- CAFs2@meta.data %>%
  select(patient.ID, CAFpopulations) %>%
  group_by(patient.ID, CAFpopulations) %>% 
  summarise(Nb = n(), .groups = "drop") %>%
    
  complete(
    patient.ID,
    CAFpopulations,
    fill = list(Nb = 0)
  ) %>%
  
  group_by(patient.ID) %>%
  mutate(
    Count = sum(Nb),
    Percent = ifelse(Count == 0, 0, Nb / Count * 100)
  ) %>%
  ungroup()

PatientData <- unique(select(Tumor@meta.data, c(patient.ID, MS.status, iCMS, CMS, KRAS, BRAF, APC, TP53, PIK3CA, Sidedness, Group.Stage, sample.origin, Site)))
PatientPlots <- left_join(PatientPlots, PatientData, by = "patient.ID")

```

#CAF subtype proportions across all patients
```{r}

CAFproportions <- PatientPlots %>%
    filter(!str_detect(sample.origin, "NA")) %>%
    group_by(sample.origin, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(sample.origin) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFproportions %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(sample.origin, "NA")) %>%
  ggplot(., aes(x = sample.origin, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

```

#CAF subtype distribution across CMSs
```{r}

CAFCMS <- PatientPlots %>%
    filter(!str_detect(CMS, "NA")) %>%
    group_by(CMS, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(CMS) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFCMS %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(CMS, "NA")) %>%
  mutate(CMS = factor(CMS, levels= c("CMS1", "CMS2", "CMS3", "CMS4")))%>%
  ggplot(., aes(x = CMS, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()


PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  #mutate(MS.status = fct_relevel(MS.status, "MSS", "MSI-H"))%>%
  filter(!str_detect(CMS, "NA"))%>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  #geom_text(aes(label = paste(Percent,"%")), position = position_stack(vjust = 0.5), size = 2)+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ CMS, scales = "free", space = "free")

```

#preCAF2 enrichment in CMS1
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(CMS, "NA"))%>%
  mutate(CMS_group = ifelse(CMS %in% c("CMS3", "CMS2", "CMS4"), "CMSall", "CMS1")) %>%
  filter(str_detect(CAFpopulations, "preCAF2"))

PatientPlots2 %>%
  mutate(CMS_group = fct_relevel(CMS_group, "CMSall", "CMS1"))%>%
  ggplot(., aes(x = CMS_group, y = Percent, color = CMS_group ))+
  scale_color_manual(values = c("CMSall" = "#53D3F3", "CMS1" = "#0C92C6")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")

```

#preCAF1 enrichment in CMS3
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(CMS, "NA"))%>%
  mutate(CMS_group = ifelse(CMS %in% c("CMS1", "CMS2", "CMS4"), "CMSall", "CMS3")) %>%
  filter(str_detect(CAFpopulations, "preCAF1"))

PatientPlots2 %>%
  mutate(CMS_group = fct_relevel(CMS_group, "CMSall", "CMS3"))%>%
  ggplot(., aes(x = CMS_group, y = Percent, color = CMS_group ))+
  scale_color_manual(values = c("CMSall" = "#FD9AFC", "CMS3" = "#F70DD1")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#mCAF enrichment in CMS4
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(CMS, "NA"))%>%
  mutate(CMS_group = ifelse(CMS %in% c("CMS1", "CMS2", "CMS3"), "CMSall", "CMS4")) %>%
  filter(str_detect(CAFpopulations, "mCAF"))

PatientPlots2 %>%
  mutate(CMS_group = fct_relevel(CMS_group, "CMSall", "CMS4"))%>%
  ggplot(., aes(x = CMS_group, y = Percent, color = CMS_group ))+
  scale_color_manual(values = c("CMSall" = "#AD7FE9", "CMS4" = "#740DE2")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#CAF subtype distribution across iCMSs
```{r}

CAFiCMS <- PatientPlots %>%
    filter(!str_detect(iCMS, "NA")) %>%
    group_by(iCMS, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(iCMS) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFiCMS %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(iCMS, "NA")) %>%
  mutate(iCMS = factor(iCMS, levels= c("iCMS2", "iCMS3")))%>%
  ggplot(., aes(x = iCMS, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ iCMS, scales = "free", space = "free")

```

#preCAF1 enrichment in iCMS2
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(iCMS, "NA"))%>%
  filter(str_detect(CAFpopulations, "preCAF1"))

PatientPlots2 %>%
  mutate(iCMS = fct_relevel(iCMS, "iCMS2", "iCMS3"))%>%
  ggplot(., aes(x = iCMS, y = Percent, color = iCMS ))+
  scale_color_manual(values = c("iCMS2" = "#F70DD1", "iCMS3" = "#FD9AFC")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#iCAF enrichment in iCMS3
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(iCMS, "NA"))%>%
  filter(str_detect(CAFpopulations, "iCAF"))

PatientPlots2 %>%
  mutate(iCMS = fct_relevel(iCMS, "iCMS2", "iCMS3"))%>%
  ggplot(., aes(x = iCMS, y = Percent, color = iCMS ))+
  scale_color_manual(values = c("iCMS2" = "#FF4BC0", "iCMS3" = "#DC043E")) +
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#CAF subtype distribution across tumor stages
```{r}

CAFStage <- PatientPlots %>%
    filter(!str_detect(Group.Stage, "NA")) %>%
    group_by(Group.Stage, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(Group.Stage) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFStage %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(Group.Stage, "NA")) %>%
  ggplot(., aes(x = Group.Stage, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ Group.Stage, scales = "free", space = "free")

```

#iCAF enrichment in early and late-stage CRC
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(Group.Stage, "NA"))%>%
  mutate(Stage = ifelse(Group.Stage %in% c("1", "2"), "1/2", "3/4")) %>%
  filter(str_detect(CAFpopulations, "iCAF"))

PatientPlots2 %>%
  mutate(Stage = factor(Stage, levels= c("1/2", "3/4")))%>%
  ggplot(., aes(x = Stage, y = Percent, color = Stage))+
  scale_color_manual(values = c("1/2" = "#DC043E", "3/4" = "#FF4BC0")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#CAF subtype distribution between right- and left-sided tumors
```{r}

CAFSide <- PatientPlots %>%
    filter(!str_detect(Sidedness, "NA")) %>%
    group_by(Sidedness, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(Sidedness) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFSide %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(Sidedness, "NA")) %>%
  mutate(Sidedness = factor(Sidedness, levels= c("R", "L")))%>%
  ggplot(., aes(x = Sidedness, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(Sidedness = factor(Sidedness, levels= c("R", "L")))%>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ Sidedness, scales = "free", space = "free")

```

#CAF subtype distribution across proximo-distal location
```{r}

CAFSide2 <- Stroma@meta.data %>%
  select(patient.ID, populations, sample.origin, Site) %>%
  filter(sample.origin == "Tumor") %>%
  filter(populations %in% c("preCAF1", "preCAF2", "iCAF", "mCAF")) %>%
  
  ## ---------- NEW: collapse individual sites into 3 regions ----------
  mutate(
    Region = case_when(
      Site %in% c(
        "Caecum", "Ascending colon", "Hepatic Flexure", "Distal Ascending colon", "Transverse colon"
      )                            ~ "Proximal colon",
      Site %in% c(
        "Descending colon", "Distal Descending colon",
        "Sigmoid colon", "Distal Sigmoid colon", "Rectosigmoid"
      )                            ~ "Distal colon",
      Site %in% c(
        "Upper rectum", "Mid-rectum", "Low rectum", "Rectum"
      )                            ~ "Rectum",
      TRUE                         ~ NA_character_   # catch any stray value
    )
  ) %>%
  filter(!is.na(Region)) %>%      # drop stray/unknown sites if any
  ## -------------------------------------------------------------------

  group_by(Region, populations) %>%  # ← use Region instead of Site
  summarise(Nb = n(), .groups = "drop_last") %>%
  mutate(Total   = sum(Nb),
         Percent = round(Nb / Total * 100, 2)) %>%
  ungroup()

# set the x-axis order you prefer
region_order <- c("Proximal colon", "Distal colon", "Rectum")
CAFSide2$Region <- factor(CAFSide2$Region, levels = region_order)

CAFSide2 %>%
  mutate(populations = factor(populations,
                              levels = c("preCAF1", "preCAF2", "iCAF", "mCAF"))) %>%
  ggplot(aes(x = Region, y = Percent, fill = populations)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = ifelse(Percent >= 5, paste0(Percent, "%"), "")),  # Only label >5% for clarity
            position = position_stack(vjust = 0.5),
            color = "white", size = 3) +
  scale_fill_manual(values = c(
    preCAF1 = "#FD9AFC",
    preCAF2 = "#53D3F3",
    iCAF    = "#FD3369",
    mCAF    = "#7012E5"
  )) +
  xlab("Tumor region") +
  ylab("Percent composition of CAFs") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

#preCAF1 enrichment in rectal tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(Site, "NA"))%>%
  mutate(
    Region = case_when(
      Site %in% c(
        "Caecum", "Ascending colon", "Hepatic Flexure", "Distal Ascending colon", "Transverse colon"
      )                            ~ "Proximal colon",
      Site %in% c(
        "Descending colon", "Distal Descending colon",
        "Sigmoid colon", "Distal Sigmoid colon", "Rectosigmoid"
      )                            ~ "Distal colon",
      Site %in% c(
        "Upper rectum", "Mid-rectum", "Low rectum", "Rectum"
      )                            ~ "Rectum",
      TRUE                         ~ NA_character_   # catch any stray value
    )
  ) %>%  filter(str_detect(CAFpopulations, "preCAF1"))

comparisons <- list(
  c("Proximal colon", "Rectum"),
  c("Distal colon", "Rectum"))

PatientPlots2 %>%
  mutate(Region = fct_relevel(Region, "Proximal colon", "Distal colon", "Rectum"))%>%
  ggplot(., aes(x = Region, y = Percent, color = Region ))+
  scale_color_manual(values = c("Proximal colon" = "#FD9AFC", "Distal colon" = "#FD9AFC", "Rectum" = "#F70DD1")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", comparisons = comparisons, label = "p.format") 

```

#CAF subtype distribution between MSS and MSI-H patients
```{r}

CAFMS <- PatientPlots %>%
    filter(!str_detect(MS.status, "NA")) %>%
    group_by(MS.status, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(MS.status) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFMS %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(MS.status, "NA")) %>%
  mutate(MS.status = factor(MS.status, levels= c("MSS", "MSI-H")))%>%
  ggplot(., aes(x = MS.status, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(MS.status = factor(MS.status, levels= c("MSS", "MSI-H")))%>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ MS.status, scales = "free", space = "free")

```

#preCAF1 enrichment in MSS tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(MS.status, "NA"))%>%
  filter(str_detect(CAFpopulations, "preCAF1"))

PatientPlots2 %>%
  mutate(MS.status = factor(MS.status, levels= c("MSS", "MSI-H")))%>%
  ggplot(., aes(x = MS.status, y = Percent, color = MS.status))+
  scale_color_manual(values = c("MSS" = "#F70DD1", "MSI-H" = "#FD9AFC")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#CAF subtype distribution between APC wild-type and mutant tumors
```{r}

CAFAPC <- PatientPlots %>%
    filter(!str_detect(APC, "NA")) %>%
    group_by(APC, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(APC) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFAPC %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(APC, "NA")) %>%
  mutate(APC = factor(APC, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = APC, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(APC = factor(APC, levels= c("wt", "mut")))%>%
   filter(!str_detect(APC, "NA")) %>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ APC, scales = "free", space = "free")

```

#CAF subtype distribution between KRAS wild-type and mutant tumors
```{r}

CAFKRAS <- PatientPlots %>%
    filter(!str_detect(KRAS, "NA")) %>%
    group_by(KRAS, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(KRAS) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFKRAS %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(KRAS, "NA")) %>%
  mutate(KRAS = factor(KRAS, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = KRAS, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(KRAS = factor(KRAS, levels= c("wt", "mut")))%>%
   filter(!str_detect(KRAS, "NA")) %>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ KRAS, scales = "free", space = "free")

```

#mCAF enrichment in KRAS mutant tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(KRAS, "NA"))%>%
  filter(str_detect(CAFpopulations, "mCAF"))

PatientPlots2 %>%
  mutate(KRAS = factor(KRAS, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = KRAS, y = Percent, color = KRAS))+
  scale_color_manual(values = c("wt" = "#AD7FE9", "mut" = "#740DE2")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#CAF subtype distribution between early- and late-stage KRAS wild-type and mutant tumors
```{r}

agg_data <- PatientPlots %>%
  filter(!str_detect(KRAS, "NA")) %>%
  mutate(
    StageGroup = case_when(
      Group.Stage %in% c("1", "2") ~ "1/2",
      Group.Stage %in% c("3", "4") ~ "3/4",
      TRUE                         ~ NA_character_
    )
  ) %>%
  filter(!is.na(StageGroup)) %>%
  group_by(KRAS, StageGroup, CAFpopulations) %>%
  summarize(Percent = sum(Percent), .groups = "drop") %>%
  group_by(KRAS, StageGroup) %>%
  mutate(
    Percent_norm = Percent / sum(Percent) * 100,
    category_reordered = factor(CAFpopulations, levels = c("preCAF1", "preCAF2", "mCAF", "iCAF"))
  ) %>%
  ungroup() %>%
  mutate(KRAS = fct_relevel(KRAS, "wt", "mut"))  # Make sure this is here

ggplot(agg_data, aes(x = KRAS, y = Percent_norm, fill = category_reordered)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", Percent_norm)), 
            position = position_stack(vjust = 0.5), size = 3, color = "white") +
  scale_fill_manual(values = c(
    "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3",
    "iCAF" = "#FD3369", "mCAF" = "#7012E5"
  )) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid(~ StageGroup, scales = "free", space = "free")

```

#mCAF enrichment in late-stage KRAS mutant tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(KRAS, "NA"))%>%
  filter(str_detect(CAFpopulations, "mCAF"))

PatientPlots2 %>%
  mutate(KRAS = factor(KRAS, levels= c("wt", "mut")))%>%
  mutate(
    StageGroup = case_when(
      Group.Stage %in% c("1", "2") ~ "1/2",
      Group.Stage %in% c("3", "4") ~ "3/4",
      TRUE                         ~ NA_character_
    )
  ) %>%
  ggplot(., aes(x = KRAS, y = Percent, color = KRAS))+
  scale_color_manual(values = c("wt" = "#AD7FE9", "mut" = "#740DE2")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  facet_wrap( ~ StageGroup)+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#CAF subtype distribution between TP53 wild-type and mutant tumors
```{r}

CAFTP53 <- PatientPlots %>%
    filter(!str_detect(TP53, "NA")) %>%
    group_by(TP53, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(TP53) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFTP53 %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(TP53, "NA")) %>%
  mutate(TP53 = factor(TP53, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = TP53, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(TP53 = factor(TP53, levels= c("wt", "mut")))%>%
   filter(!str_detect(TP53, "NA")) %>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ TP53, scales = "free", space = "free")

```

#preCAF1 enrichment in TP53 mutant tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(TP53, "NA"))%>%
  filter(str_detect(CAFpopulations, "preCAF1"))

PatientPlots2 %>%
  mutate(TP53 = factor(TP53, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = TP53, y = Percent, color = TP53))+
  scale_color_manual(values = c("wt" = "#FD9AFC", "mut" = "#F70DD1")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value


```

#CAF subtype distribution between BRAF wild-type and mutant tumors
```{r}

CAFBRAF <- PatientPlots %>%
    filter(!str_detect(BRAF, "NA")) %>%
    group_by(BRAF, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(BRAF) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFBRAF %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>%
  filter(!str_detect(BRAF, "NA")) %>%
  mutate(BRAF = factor(BRAF, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = BRAF, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(BRAF = factor(BRAF, levels= c("wt", "mut")))%>%
   filter(!str_detect(BRAF, "NA")) %>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ BRAF, scales = "free", space = "free")

```

#preCAF2 enrichment in BRAF wt tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(BRAF, "NA"))%>%
  filter(str_detect(CAFpopulations, "preCAF2"))

PatientPlots2 %>%
  mutate(BRAF = factor(BRAF, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = BRAF, y = Percent, color = BRAF))+
  scale_color_manual(values = c("wt" = "#0C92C6", "mut" = "#53D3F3")) +
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value


```

#CAF subtype distribution between CAFPIK3CA wild-type and mutant tumors
```{r}

CAFPIK3CA <- PatientPlots %>%
    filter(!str_detect(PIK3CA, "NA")) %>%
    group_by(PIK3CA, CAFpopulations) %>%
    summarise(
        AvgPercent = mean(Percent, na.rm = TRUE),
        .groups = 'drop'
    ) %>%
    group_by(PIK3CA) %>%
    mutate(AvgPercent = AvgPercent / sum(AvgPercent) * 100) %>%
    ungroup() %>%
    mutate(AvgPercent = round(AvgPercent, 1))

CAFPIK3CA %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  filter(!str_detect(PIK3CA, "NA")) %>%
  mutate(PIK3CA = factor(PIK3CA, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = PIK3CA, y = AvgPercent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = paste(AvgPercent,"%")), position = position_stack(vjust = 0.5))+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()

PatientPlots %>%
  mutate(category_reordered = factor(CAFpopulations, levels =  c("preCAF1", "iCAF","preCAF2", "mCAF"))) %>% 
  mutate(PIK3CA = factor(PIK3CA, levels= c("wt", "mut")))%>%
   filter(!str_detect(PIK3CA, "NA")) %>%
  ggplot(., aes(x = patient.ID, y = Percent, fill = category_reordered ))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c( "preCAF1" = "#FD9AFC", "preCAF2" = "#53D3F3", "iCAF" = "#FD3369", "mCAF" = "#AD7FE9")) + 
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  facet_grid( ~ PIK3CA, scales = "free", space = "free")

```

#preCAF1 enrichment in PIK3CA wt tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(PIK3CA, "NA"))%>%
  filter(str_detect(CAFpopulations, "preCAF1"))

PatientPlots2 %>%
  mutate(PIK3CA = factor(PIK3CA, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = PIK3CA, y = Percent, color = PIK3CA))+
  scale_color_manual(values = c("wt" = "#F70DD1", "mut" = "#FD9AFC")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#iCAF enrichment in PIK3CA mutant tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(PIK3CA, "NA"))%>%
  filter(str_detect(CAFpopulations, "iCAF"))

PatientPlots2 %>%
  mutate(PIK3CA = factor(PIK3CA, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = PIK3CA, y = Percent, color = PIK3CA))+
  scale_color_manual(values = c("wt" = "#FF4BC0", "mut" = "#DC043E")) + 
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```

#mCAF enrichment in PIK3CA mutant tumors
```{r}

PatientPlots2 <- PatientPlots %>%
  filter(!str_detect(PIK3CA, "NA"))%>%
  filter(str_detect(CAFpopulations, "mCAF"))

PatientPlots2 %>%
  mutate(PIK3CA = factor(PIK3CA, levels= c("wt", "mut")))%>%
  ggplot(., aes(x = PIK3CA, y = Percent, color = PIK3CA))+
  scale_color_manual(values = c("wt" = "#AD7FE9", "mut" = "#740DE2")) +
  geom_boxplot()+
  geom_point(position=position_dodge(width = 0.75))+
  theme_minimal()+
  expand_limits(y = c(0:100))+
  stat_compare_means(method = "wilcox.test", label = "p.format")  # Add p-value

```




