---
title: "Code2_mouse_integration_scRNAseq"
output: html_document
date: "2025-07-14"
---

#load libraries
```{r}

library(Seurat)
library(tidyverse)
library(patchwork)
library(biomaRt)
library(RColorBrewer)

```

#create Seurat objects
```{r}

RoulisC1A2 <- readRDS("/dir/RoulisC1A.rds")
RoulisC1A2@meta.data[["orig.ident"]] <- "Roulis"

VlnPlot(RoulisC1A2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC1A2, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

RoulisC1A2 <- subset(RoulisC1A2, subset =
                       nFeature_RNA > 150 & nFeature_RNA < 1500 & 
                       nCount_RNA > 150 & nCount_RNA < 3000 & 
                       percent.mt < 15 & percent.rb < 25)

VlnPlot(RoulisC1A2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC1A2, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

RoulisC2A2 <- readRDS("/dir/RoulisC2A.rds")
RoulisC2A2@meta.data[["orig.ident"]] <- "Roulis"


VlnPlot(RoulisC2A2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC2A2, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

RoulisC2A2 <- subset(RoulisC2A2, subset =
                       nFeature_RNA > 150 & nFeature_RNA < 1500 & 
                       nCount_RNA > 150 & nCount_RNA < 3000 & 
                       percent.mt < 15 & percent.rb < 25)

VlnPlot(RoulisC2A2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC2A2, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

Jasso <- readRDS("/dir/Jasso.rds")
Jasso@meta.data[["orig.ident"]] <- "Jasso"

VlnPlot(Jasso, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Jasso, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

Jasso <- subset(Jasso, subset =
                       nFeature_RNA > 600 & nFeature_RNA < 2500 & 
                       nCount_RNA > 1000 & nCount_RNA < 6000 & 
                       percent.mt < 5 & percent.rb < 25)

VlnPlot(Jasso, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Jasso, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

Kinchen <- readRDS("/dir/Kinchen.rds")
Kinchen@meta.data[["orig.ident"]] <- "Kinchen"

VlnPlot(Kinchen, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Kinchen, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")


Kinchen <- subset(Kinchen, subset =
                       nFeature_RNA > 500 & nFeature_RNA < 3500 & 
                       nCount_RNA > 2000 & nCount_RNA < 3500 & 
                       percent.mt < 2 & percent.rb < 12)

VlnPlot(Kinchen, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Kinchen, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

Brugger <- readRDS("/dir/Brugger.rds")
Brugger@meta.data[["orig.ident"]] <- "Brügger"

VlnPlot(Brugger, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Brugger, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

Brugger <- subset(Brugger, subset =
                       nFeature_RNA > 1500 & nFeature_RNA < 4000 & 
                       nCount_RNA > 7500 & nCount_RNA < 10500 & 
                       percent.mt < 20 & percent.rb < 30)

VlnPlot(Brugger, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Brugger, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

ThisStudy <- readRDS("/dir/ThisStudy.rds")
ThisStudy@meta.data[["orig.ident"]] <- "ThisStudy"

VlnPlot(ThisStudy, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(ThisStudy, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

ThisStudy <- subset(ThisStudy, subset = nFeature_RNA > 1000 & nFeature_RNA < 4000
                         & nCount_RNA < 13000 & nCount_RNA > 1000
                         & percent.mt < 6
                         & percent.rb < 25 & percent.rb > 3)

VlnPlot(ThisStudy, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)

VlnPlot(ThisStudy, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(ThisStudy, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

#normalize and integrate data, run PCA
```{r}

datalist <- list(RoulisC1A2, RoulisC2A2 , JassoH2O, KinchenHC, Brugger, Koliaraki)

convertHumanGeneList <- function(x){
  require("biomaRt")
  human = useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://dec2021.archive.ensembl.org/")
  mouse = useEnsembl("ensembl", dataset = "mmusculus_gene_ensembl", host = "https://dec2021.archive.ensembl.org/")
  genes = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = x , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)
  humanx <- unique(genes[, 2])
  print(head(humanx))
  return(humanx)
}

m.s.genes <- convertHumanGeneList(cc.genes$s.genes)
m.g2m.genes <- convertHumanGeneList(cc.genes$g2m.genes)

for (i in 1:length(datalist)) {
    datalist[[i]] <- SCTransform(datalist[[i]])
    datalist[[i]] <- CellCycleScoring(datalist[[i]], g2m.features = m.g2m.genes, s.features= m.s.genes)
    datalist[[i]] <- SCTransform(datalist[[i]], vars.to.regress = c( "percent.mt", "percent.rb", "S.Score", "G2M.Score"))
    }

features <- SelectIntegrationFeatures(object.list = datalist, nfeatures = 3000)

datalist <- PrepSCTIntegration(object.list = datalist, anchor.features = features)

datalist <- lapply(X = datalist, FUN = RunPCA, features = features)

anchors <- FindIntegrationAnchors(object.list = datalist, normalization.method = "SCT",
                                  anchor.features = features, reduction = "rpca", k.anchor = 7)

DataIntegrated <- IntegrateData(anchorset = anchors, normalization.method = "SCT")




DefaultAssay(DataIntegrated) <- "integrated"

DataIntegrated <- RunPCA(DataIntegrated, verbose = FALSE)
ElbowPlot(DataIntegrated)
DataIntegrated <- RunUMAP(DataIntegrated, reduction = "pca", dims = 1:30)

DataIntegrated <- FindNeighbors(DataIntegrated, reduction = "pca", dims = 1:30)
DataIntegrated <- FindClusters(DataIntegrated, resolution = 0.7)

DefaultAssay(DataIntegrated) <- "RNA"
DataIntegrated <- NormalizeData(DataIntegrated, verbose = F)
DataIntegrated <- FindVariableFeatures(DataIntegrated, selection.method = "vst", nfeatures = 3000, verbose = FALSE)
DataIntegrated <- ScaleData(DataIntegrated, verbose = FALSE)

DimPlot(DataIntegrated, reduction = "umap", label = T, label.size = 4)

```

#split data by original dataset
```{r}
DimPlot(DataIntegrated, reduction = "umap", label = F, split.by = "orig.ident")
```

#plot cell type marker
```{r}

DefaultAssay(DataIntegrated) <- "RNA"

FeaturePlot(DataIntegrated, c( "Pdgfra", "Acta2", "Epcam", "Lgr5", "Muc2",  "Ptprc", "Pecam1", "Lyve1", "Notch3", "Ano1", "Kit", "S100b", "Tubb3" ), order = T)

```

#check cluster quality
```{r}

FeaturePlot(DataIntegrated, c("nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rb"), order =T)

```

#filter out immune and poor quality cells
```{r}

DataIntegratedSub <- subset(DataIntegrated, idents = c(0,1,2,3,4,5,6,7,8,9,10, 12,13,14,15, 17))

DefaultAssay(DataIntegratedSub) <- "integrated"

DataIntegratedSub <- RunPCA(DataIntegratedSub, verbose = FALSE)
DataIntegratedSub <- RunUMAP(DataIntegratedSub, reduction = "pca", dims = 1:15)

DataIntegratedSub <- FindNeighbors(DataIntegratedSub, reduction = "pca", dims = 1:15)
DataIntegratedSub <- FindClusters(DataIntegratedSub, resolution = 0.6)


DimPlot(DataIntegratedSub, reduction = "umap", label = T, pt.size = 0.1)

```

#find marker genes
```{r}

Markers <- FindAllMarkers(DataIntegratedSub, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.25, assay = "RNA")

top10 <- Markers %>%
  group_by(cluster)%>%
  top_n(n = 10, wt = avg_log2FC)

DefaultAssay(DataIntegratedSub) <- "RNA"
VlnPlot(DataIntegratedSub, features =  top10$gene, pt.size = 0)

```

#name clusters
```{r}

ClusterIDs <- c( "dist_bottom_MAF", 
                 "SAF", "bottom_SEMF", "prox_bottom_MAF","LP_SMC/MM_SMC",  "top_SEMF", 
                 "VEC", "colonocyte", "top_MAF", "LEC","Goblet",
                 "glia",  "peri", "MIF", "EC+peri")

names(ClusterIDs) <- levels(DataIntegratedSub)
DataIntegrated2 <- RenameIdents(DataIntegratedSub, ClusterIDs)

DataIntegrated2@active.ident <- factor(DataIntegrated2@active.ident,
                                       levels = c( "top_SEMF", "bottom_SEMF", "top_MAFF",  "prox_bottom_MAF", "dist_bottom_MAF",
                                                  "SEF", "MIF","LP_SMC/MM_SMC",
                                                  "peri","EC+peri", "VEC", "LEC", "glia",
                                                  "colonocyte","Goblet"))

DimPlot(DataIntegrated2, reduction = "umap", label = F, pt.size = 0.5,
       cols = c("top_SEMF" = "#FCED1D", "bottom_SEMF" = "#F6A937",
                "top_MAF" = "#1DE0FC", "prox_bottom_MAF" = "#1D7BFC", "dist_bottom_MAF" = "#182FAF",
                'SAF' = '#129B93',  "MIF" = "#035C56", 
                "LP_SMC/MM_SMC" = "#C51B7C", 
                "peri" = "#6E5F57", "EC+peri" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E",
                "glia" = "#C2D9F0", "colonocyte" = "#C3936D","Goblet" = "#734724")) & NoAxes()

saveRDS(DataIntegrated2, "/dir/Mm_colon_integration.rds")

```

#split data by original dataset
```{r}

DimPlot(DataIntegrated2, reduction = "umap", label = F, pt.size = 0.1, split.by = "orig.ident",
       cols = c("top_SEMF" = "#FCED1D", "bottom_SEMF" = "#F6A937",
                "top_MAF" = "#1DE0FC", "prox_bottom_MAF" = "#1D7BFC", "dist_bottom_MAF" = "#182FAF",
                'SAF' = '#129B93',  "MIF" = "#035C56", 
                "LP_SMC/MM_SMC" = "#C51B7C", 
                "peri" = "#6E5F57", "EC+peri" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E",
                "glia" = "#C2D9F0", "colonocyte" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#split data by sampling site
```{r}

DimPlot(DataIntegrated2, split.by = "Site", pt.size = 0.2,
       cols = c("top_SEMF" = "#FCED1D", "bottom_SEMF" = "#F6A937",
                "top_MAF" = "#1DE0FC", "prox_bottom_MAF" = "#1D7BFC", "dist_bottom_MAF" = "#182FAF",
                'SAF' = '#129B93',  "MIF" = "#035C56", 
                "LP_SMC/MM_SMC" = "#C51B7C", 
                "peri" = "#6E5F57", "EC+peri" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E",
                "glia" = "#C2D9F0", "colonocyte" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#find markers between fibroblast populations
```{r}

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "top_SEMF",
                       ident.2 = c("bottom_SEMF", "top_MAF", "prox_bottom_MAF", "dist_bottom_MAF", "SAF", "MIF", "LP_SMC/MM_SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "top_SEMF",
                       ident.2 = c("bottom_SEMF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")


Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "bottom_SEMF",
                       ident.2 = c("top_SEMF", "top_MAF", "prox_bottom_MAF", "dist_bottom_MAF", "SAF", "MIF", "LP_SMC/MM_SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "bottom_SEMF",
                       ident.2 = c("top_SEMF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")




Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "top_MAF",
                       ident.2 = c("top_SEMF", "bottom_SEMF", "prox_bottom_MAF", "dist_bottom_MAF", "SAF", "MIF", "LP_SMC/MM_SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.35, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "top_MAF",
                       ident.2 = c( "prox_bottom_MAF", "dist_bottom_MAF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.35, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers1 <- FindMarkers(DataIntegrated2,
                       ident.1 = "prox_bottom_mucF",
                       ident.2 = c("dist_bottom_mucF"),
                       logfc.threshold = 0.3, min.diff.pct = 0.3, min.pct = 0.3, only.pos = T, assay ="RNA")


Markers2 <- FindMarkers(DataIntegrated2,
                       ident.1 = "dist_bottom_mucF",
                       ident.2 = c("prox_bottom_mucF"),
                       logfc.threshold = 0.3, min.diff.pct = 0.3, min.pct = 0.3, only.pos = T, assay ="RNA")


Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "dist_bottom_MAF",
                       ident.2 = c("top_SEMF", "top_MAF", "prox_bottom_MAF", "bottom_SEMF", "SAF", "MIF", "LP_SMC/MM_SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.4, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "dist_bottom_MAF",
                       ident.2 = c( "top_MAF", "prox_bottom_MAF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.3, min.pct = 0.5, only.pos = T, assay ="RNA")




Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "SAF",
                       ident.2 = c("top_SEMF", "top_MAF", "prox_bottom_MAF", "dist_bottom_MAF","bottom_SEMF", "MIF", "LP_SMC/MM_SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "MIF",
                       ident.2 = c("top_SEMF", "top_MAF", "prox_bottom_MAF", "dist_bottom_MAF", "SAF", "bottom_SEMF", "LP_SMC/MM_SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")


Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "LP_SMC/MM_SMC",
                       ident.2 = c("top_SEMF", "top_MAF", "prox_bottom_MAF", "dist_bottom_MAF", "SAF", "bottom_SEMF", "MIF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.55, min.pct = 0.5, only.pos = T, assay ="RNA")


VlnPlot(DataIntegrated2, rownames(Markers), pt.size = 0)

```

#plot validated and new fibroblast subpopulation markers
```{r}

DataIntegrated2@active.ident <- factor(DataIntegrated2@active.ident,
                                       levels = c("Goblet","colonocyte", "glia","lymph_endoth","vasc_endoth", "endoth+peri","peri", 
                                                  "LP_SMC/MM_SMC","MIF", "SAF", 
                                                 "dist_bottom_MAF", "prox_bottom_MAF","top_MAF",  "bottom_SEMF", "top_SEMF"))

genes <- c("Pdgfra", "Foxl1", "F3", "Procr", "Bmp5", "Bmp7" , "Wnt5a", "Acta2", "Wnt4", "Fgfr2", "Igfbp5", "Cd34", "Cd81", "Wnt2b", "Grem1", "Grem2", "Rspo3", "Lama5", "Wif1", "Pi16", "Kcnn3")

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size", col.min = 0)+
  theme(axis.text=element_text(angle=90))

```

#markers for other cell types except for fibroblasts
```{r}

genes <- c( "Acta2",  "Lmod1",  "Cnn1", "Tagln", "Myh11", "Myl9",  "Mylk",  "Actg2","Des",
            "Pdgfrb", "Cspg4" , "Rgs4", "Vtn", "Notch3",
            "Flt1","Pecam1", 
            "Lyve1",  "Flt4", "Prox1",
            "S100b", "Gfap", "Uchl1",
            "Epcam", "Aldob", "Fabp2", "Muc2")

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size")+
    theme(axis.text=element_text(angle=90))

```

#signature plot for stem cell niche factors
```{r}

genelist1 <- list(c("Grem1", "Grem2", "Rspo3"))

DataIntegrated2 <- AddModuleScore(object = DataIntegrated2, features = genelist1, name = "genelist1")

FeaturePlot(object = DataIntegrated2, features = "genelist1", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#signature plot for differentiation niche factors
```{r}

genelist2 <- list(c("Bmp5", "Bmp7", "Wnt5a"))

DataIntegrated2 <- AddModuleScore(object = DataIntegrated2, features = genelist2, name = "genelist2")

FeaturePlot(object = DataIntegrated2, features = "genelist2", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#mouse markers for all subpopulations
```{r}

genes <- c( "Bmp7",  "Fgf9", "Col4a5", "Col4a6","Agt", "Nrg1", "Glp2r",   "Trpa1",
            "Sox6","Cxcl14", "Ednrb", "Bmp5","Insc",   "Lama5",
            "Fgf7", "Cd38", "Ccdc3", 
            "Aldh1a1",   "Ogn","Thbd", "Gnai1", "Bmp3", "Wif1",
            
            "Vcam1", "Ccl8", "Hapln1", "Edil3",   "Stmn2",
            "Cxcl10" ,"Tnfaip2" ,  "Hcar2", "Prrx2",
            "Igfbp3", "Cxcl16", "Serpina3n","Igfpb5",  "Fgf10",  "Rad51b", "Scube1","Lepr",
            "Osr1",  "Tnfsf13b", "Eln",   "Cd248",
            
            "Medag", "Pi16","Pcolce2", "Cilp", "Osr2",
             "C3", "Has1", "Adamts5", "Ar", "Ackr4",
             "Scara5", "Il33", "Dpep1", "Gfpt2",
             
             "Kcnn3",  "P2ry1", "Ramp1", "Fibin", "Smoc2", "Pid1", "Thbs2", "Jam2", "Fxyd6", "Mgp",
             "Ank1", "Gdf10", "Ngf", "Cox8b",

             "Hhip",
             "Acta2",  "Lmod1",  "Cnn1","Tagln", "Myh11", "Myl9",  "Mylk",  "Actg2","Des")

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size")+
    theme(axis.text=element_text(angle=90))

```

#human markers that do not recapitulate in mouse
```{r}

genes <- c( "Vstm2a", "Nsg1", "Enho",
            "Hsd17b2", "Frzb",  "Prom1", "F2r",  "Pdgfd", "Dmkn",
            "Adam28", "Adamdec1",  "Abca8", "Sftpf",
            "F10", "Wisp2",  "Fbln2", "Sfrp1","Sfrp2",   "Igf1", "Slpi",
            "Thbs4", "Fndc1", "Adamtsl3", "C7")

DotPlot(DataIntegrated2Hs, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size", col.min = 0)+
    theme(axis.text=element_text(angle=90))

```

#top and bottom markers
```{r}

VlnPlot(DataIntegrated2, c( "Wnt4", "Grem1"), pt.size = 0,
               cols = c("top_SEMF" = "#FCED1D", "bottom_SEMF" = "#F6A937",
                "top_MAF" = "#1DE0FC", "prox_bottom_MAF" = "#1D7BFC", "dist_bottom_MAF" = "#182FAF",
                'SAF' = '#129B93',  "MIF" = "#035C56", 
                "LP_SMC/MM_SMC" = "#C51B7C", 
                "peri" = "#6E5F57", "EC+peri" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E",
                "glia" = "#C2D9F0", "colonocyte" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#proximal distal markers
```{r}

VlnPlot(DataIntegrated2, c( "Tnc", "Cd55"), pt.size = 0,
              cols = c("top_SEMF" = "#FCED1D", "bottom_SEMF" = "#F6A937",
                "top_MAF" = "#1DE0FC", "prox_bottom_MAF" = "#1D7BFC", "dist_bottom_MAF" = "#182FAF",
                'SAF' = '#129B93',  "MIF" = "#035C56", 
                "LP_SMC/MM_SMC" = "#C51B7C", 
                "peri" = "#6E5F57", "EC+peri" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E",
                "glia" = "#C2D9F0", "colonocyte" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#top and bottom semf markers
```{r}

DefaultAssay(DataIntegrated2) <- "RNA"

FeaturePlot(DataIntegrated2, c( "Lama5", "Lgr5", "Adamts18", "Wif1", "Bmp3", "Acta2"), order =T)

```

#subset and recluster SMCs
```{r}

DataIntegrated2Sub <- subset(DataIntegrated2, idents = c("LP_SMC/MM_SMC"))

DefaultAssay(DataIntegrated2Sub) <- "integrated"

DataIntegrated2Sub <- RunPCA(DataIntegrated2Sub, verbose = FALSE)
DataIntegrated2Sub <- RunUMAP(DataIntegrated2Sub, reduction = "pca", dims = 1:10)

DataIntegrated2Sub <- FindNeighbors(DataIntegrated2Sub, reduction = "pca", dims = 1:10)
DataIntegrated2Sub <- FindClusters(DataIntegrated2Sub, resolution = 0.075)

DimPlot(DataIntegrated2Sub, reduction = "umap", label = T)

```

#find SMC markers
```{r}

DefaultAssay(DataIntegrated2Sub) <- "RNA"

Markers <- FindAllMarkers(DataIntegrated2Sub, only.pos = TRUE, min.pct = 0.4, logfc.threshold = 0.4, min.diff.pct = 0.4, assay = "RNA")

FeaturePlot(DataIntegrated2, rownames(Markers), order =T) & NoAxes()

```

#rename SMC clusters
```{r}

ClusterIDs <- c( "myocytes", "MM_SMC")
names(ClusterIDs) <- levels(DataIntegrated2Sub)
DataIntegrated2Sub <- RenameIdents(DataIntegrated2Sub, ClusterIDs)

DimPlot(DataIntegrated2Sub, reduction = "umap", label = F, cols = c("myocytes" = "#D54B76","MM_SMC" = "#F6AEE5"), pt.size = 0.7) & NoAxes()

```

#modify function for stacked vlnplot
```{r}
# remove the x-axis text and tick
# plot.margin to adjust the white space between each plot.
# ... pass any arguments to VlnPlot in Seurat
modify_vlnplot<- function(obj, 
                          feature, 
                          pt.size = 0, 
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  p<- VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    xlab("") + ylab(feature) + ggtitle("") + 
    theme(legend.position = "none", 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(), 
          axis.title.y = element_text(size = rel(1), angle = 0), 
          axis.text.y = element_text(size = rel(1)), 
          plot.margin = plot.margin ) 
  return(p)
}

# extract the max value of the y axis
extract_max<- function(p){
  ymax<- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}

# main function
StackedVlnPlot<- function(obj, features,
                          pt.size = 0, 
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  plot_list<- purrr::map(features, function(x) modify_vlnplot(obj = obj,feature = x, ...))
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]]<- plot_list[[length(plot_list)]] +
    theme(axis.text.x=element_text(), axis.ticks.x = element_line())
  # change the y-axis tick to only max value 
  ymaxs<- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x + 
                            scale_y_continuous(breaks = c(y)) + 
                            expand_limits(y = y))
  p<- patchwork::wrap_plots(plotlist = plot_list, ncol = 1)
  return(p)
}

```

#plot SMC markers
```{r}

DataIntegrated2Sub@active.ident <- factor(DataIntegrated2Sub@active.ident, levels = c( "MM_SMC","myocytes"))

genes <- c("Acta2", "Myh11",  "Actg2","Des", "Hhip", "Kit")

StackedVlnPlot(DataIntegrated2Sub, genes,  cols = c("myocytes" = "#D54B76","MM SMC" = "#F6AEE5"))

```

#myocyte and SMC marker plot
```{r}

genes <- c( "Hhip", "Sncaip", "Pdgfc",
           "Grem1", "Grem2", "Rspo3", "Chrdl1","Id4", "Nrtn", "Foxp2", "Ptger1", "Ccl11", "Des")

DotPlot(DataIntegrated2Sub, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size", col.min = 0)+
    theme(axis.text=element_text(angle=90))

```

#signature plot myocyte
```{r}

genelist3 <- list(c( "Hhip", "Sncaip", "Pdgfc"))

DataIntegrated2 <- AddModuleScore(object = DataIntegrated2, features = genelist3, name = "genelist3")

FeaturePlot(object = DataIntegrated2, features = "genelist3", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#signature plot MM SMC
```{r}

genelist4 <- list(c( "Grem1", "Grem2", "Rspo3", "Chrdl1","Id4", "Ntrn", "Foxp2", "Ptger1", "Ccl11", "Des"))

DataIntegrated2 <- AddModuleScore(object = DataIntegrated2, features = genelist4, name = "genelist4")

FeaturePlot(object = DataIntegrated2, features = "genelist4", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#plot top MIF markers
```{r}

genes <- c( "Kcnn3", "P2ry1", "Ramp1", "Pid1", "Thbs2", "Jam2", "Fxyd6")

DotPlot(DataIntegrated2, genes, assay = "RNA", dot.min = 0.05, scale.by = "size", col.min = 0)+
    theme(axis.text=element_text(angle=90))
```


