---
title: "Code1_colon_SI_scRNAseq"
output: html_document
date: "2025-07-14"
---

#load libraries
```{r load libraries}

library(Seurat)
library(tidyverse)
library(patchwork)

```


#COLON
#load data
```{r load data}

SeuratAddCov <- function(SeuObj, species){
  
  if(species == "mmu"){
    data <- PercentageFeatureSet(SeuObj, pattern="^mt-", col.name = "percent.mt")
    data <- PercentageFeatureSet(data, pattern="^Rp[l|s][[:digit:]]", col.name="percent.rb")  
    
  }else if(species == "hs"){
    data <- PercentageFeatureSet(SeuObj, pattern="^MT-", col.name = "percent.mt")
    data <- PercentageFeatureSet(data, pattern="^RP[L|S][[:digit:]]", col.name="percent.rb")
  }
  return(data)
}


SeuratCreateObj <- function(dir, name){
  data <- Read10X(dir)
  SeuObj <- CreateSeuratObject(data, name, min.cells = 3, min.features = 200)
  SeuObj <- SeuratAddCov(SeuObj, "mmu")
  saveRDS(SeuObj, paste0("SeuObj",name, ".rds"))
  return(SeuObj)
}


if(!file.exists("SeuObjColon.rds")){
  dir <- ("/dir")
  SeuObjColon <- SeuratCreateObj(dir, "Colon")
  
}else{
  SeuObjColon <- readRDS("/dir/SeuObjColon.rds")
  
}

```

#QC plots
```{r QC plots}

VlnPlot(SeuObjColon, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)

FeatureScatter(SeuObjColon, feature1 = "nCount_RNA", feature2 = "percent.mt")

FeatureScatter(SeuObjColon, feature1 = "nCount_RNA", feature2 = "percent.rbl")

FeatureScatter(SeuObjColon, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

```

#filter object
```{r filter object}
SeuObjColonSub <- subset(SeuObjColon, subset = nFeature_RNA > 1000 & nFeature_RNA < 4000
                         & nCount_RNA < 13000 & nCount_RNA > 1000
                         & percent.mt < 6
                         & percent.rbl < 25 & percent.rbl > 3)

VlnPlot(SeuObjColonSub, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rbl"), ncol = 4)

```

#normalize data, run PCA
```{r normalize data run PCA}

Colon <- SCTransform(SeuObjColonSub)

DefaultAssay(Colon) <- "SCT"

Colon <- RunPCA(Colon, verbose = FALSE)
ElbowPlot(Colon)
Colon <- RunUMAP(Colon, reduction = "pca", dims = 1:20)

Colon <- FindNeighbors(Colon, reduction = "pca", dims = 1:20)
Colon <- FindClusters(Colon, resolution = 0.5)


DefaultAssay(Colon) <- "RNA"
Colon <- NormalizeData(Colon, verbose = F)
Colon <- FindVariableFeatures(Colon, selection.method = "vst", nfeatures = 3000, verbose = FALSE)
Colon <- ScaleData(Colon, verbose = FALSE)

DimPlot(Colon, reduction = "umap", label = T, label.size = 4) & NoAxes()

```

#identify low quality clusters
```{r identify low quality clusters}

FeaturePlot(Colon, c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rbl"), order =T)

```

#identify cell types
```{r identify cell types}

DefaultAssay(Colon) <- "RNA"

FeaturePlot(Colon, c("Pdgfra", "Myh11", "S100b", "Pecam1", "Lyve1", "Ptprc", "Kit", "Ano1", "Tubb3"), order =T)

```

#filter out low quality and immune cells, re-cluster data
```{r filter out low quality and immune cells re-cluster data}

ColonFiltered <- subset(Colon, idents = c(0, 1, 2 ,3, 4, 5, 6, 7, 10))

DefaultAssay(ColonFiltered) <- "SCT"

ColonFiltered <- RunPCA(ColonFiltered, verbose = FALSE)
ElbowPlot(ColonFiltered)
ColonFiltered <- RunUMAP(ColonFiltered, reduction = "pca", dims = 1:15)

ColonFiltered <- FindNeighbors(ColonFiltered, reduction = "pca", dims = 1:15)
ColonFiltered <- FindClusters(ColonFiltered, resolution = 0.15)

DimPlot(ColonFiltered, label =T)

```

#plot marker genes
```{r plot marker genes}

FeaturePlot(ColonFiltered, c( "Pdgfra", "Foxl1", "Pi16",
                     "Ppi16", "Cd81", "Myh11",
                     "Wnt2b", "Grem1", "Rspo3"), order =T) & NoAxes()

```

#name clusters
```{r name clusters}

ClusterIDs <- c( "PdgfraLow_Pi16-_Cd81-_1", 
                 "PdgfraLow_Pi16-_Cd81-_2", "PdgfraHigh", "PdgfraLow_Pi16+_Cd81+", "Myh11+_SMC")


names(ClusterIDs) <- levels(ColonFiltered)
ColonFiltered <- RenameIdents(ColonFiltered, ClusterIDs)

ColonFiltered@active.ident <- factor(ColonFiltered@active.ident,
                                       levels = c( "PdgfraHigh", "PdgfraLow_Pi16-_Cd81-_1",  "PdgfraLow_Pi16-_Cd81-_2", "PdgfraLow_Pi16+_Cd81+", "Myh11+_SMC"))


DimPlot(ColonFiltered, cols = c( "PdgfraHigh" = "#F6A937",
                 "PdgfraLow_Pi16-_Cd81-_1" = "#1D7BFC", "PdgfraLow_Pi16-_Cd81-_2" = "#182FAF", 
              'PdgfraLow_Pi16+_Cd81+' = '#129B93', 
              "Myh11+_SMC" = "#C51B7C")) & NoAxes()

```

#find and plot marker genes
```{r find and plot marker genes}

Markers <- FindAllMarkers(ColonFiltered, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.25, assay = "RNA")

top10 <- Markers %>%
  group_by(cluster)%>%
  top_n(n = 10, wt = avg_log2FC)

DefaultAssay(ColonFiltered) <- "RNA"

VlnPlot(ColonFiltered, features = top10$gene, pt.size = 0, cols = c( "PdgfraHi" = "#F6A937",
                 "PdgfraLow_Cd34+_1" = "#1D7BFC", "PdgfraLow_Cd34+_2" = "#182FAF", 
              'PdgfraLow_Cd34+_Cd81+' = '#129B93', 
              "Hhip+_Myofib" = "#C51B7C"), ncol = 8) & NoAxes()

saveRDS(ColonFiltered, "/dir/Mm_Colon_this_study.rds")

```








#SMALL INTESTINE
#load data
```{r load data}

SeuratAddCov <- function(SeuObj, species){
  
  if(species == "mmu"){
    data <- PercentageFeatureSet(SeuObj, pattern="^mt-", col.name = "percent.mt")
    data <- PercentageFeatureSet(data, pattern="^Rp[l|s][[:digit:]]", col.name="percent.rb")  
    
  }else if(species == "hs"){
    data <- PercentageFeatureSet(SeuObj, pattern="^MT-", col.name = "percent.mt")
    data <- PercentageFeatureSet(data, pattern="^RP[L|S][[:digit:]]", col.name="percent.rb")
  }
  return(data)
}


SeuratCreateObj <- function(dir, name){
  data <- Read10X(dir)
  SeuObj <- CreateSeuratObject(data, name, min.cells = 3, min.features = 200)
  SeuObj <- SeuratAddCov(SeuObj, "mmu")
  saveRDS(SeuObj, paste0("SeuObj",name, ".rds"))
  return(SeuObj)
}


if(!file.exists("SeuObjSI.rds")){
  dir <- ("/dir")
  SeuObjSI <- SeuratCreateObj(dir, "SI")
  
}else{
  SeuObjSI <- readRDS("/dir/SeuObjSI.rds")
  
}

```

#QC plots
```{r QC plots}

VlnPlot(SeuObjSI, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)

FeatureScatter(SeuObjSI, feature1 = "nCount_RNA", feature2 = "percent.mt")

FeatureScatter(SeuObjSI, feature1 = "nCount_RNA", feature2 = "percent.rbl")

FeatureScatter(SeuObjSI, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

```

#filter object
```{r filter object}

SeuObjSISub <- subset(SeuObjSI, subset = nFeature_RNA > 500 & nFeature_RNA < 5000
                         & nCount_RNA < 15000 & nCount_RNA > 750
                         & percent.mt < 5.5
                         & percent.rbl < 20 & percent.rbl > 2.5)

VlnPlot(SeuObjSISub, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)

```

#normalize data, run PCA
```{r normalize data and run PCA}

SI <- SCTransform(SeuObjSISub, vars.to.regress = c("percent.mt", "percent.rbl"))


DefaultAssay(SI) <- "SCT"

SI <- RunPCA(SI, verbose = FALSE)
ElbowPlot(SI)
SI <- RunUMAP(SI, reduction = "pca", dims = 1:25)

SI <- FindNeighbors(SI, reduction = "pca", dims = 1:25)
SI <- FindClusters(SI, resolution = 0.4)

DefaultAssay(SI) <- "RNA"
SI <- NormalizeData(SI, verbose = F)
SI <- FindVariableFeatures(SI, selection.method = "vst", nfeatures = 3000, verbose = FALSE)
SI <- ScaleData(SI, verbose = FALSE)

DimPlot(SI, reduction = "umap", label = T, label.size = 4)

```

#identify low quality clusters
```{r identify low quality clusters}

FeaturePlot(SI, c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rbl"), order =T)

```

#identify cell types
```{r identify cell types}

DefaultAssay(SI) <- "RNA"

FeaturePlot(SI, c("Pdgfra", "Myh11", "S100b", "Pecam1", "Lyve1", "Ptprc", "Kit", "Ano1", "Tubb3"), order =T)

```

#filter out low quality and immune cells, re-cluster data
```{r filter out low quality and immune cells, re-cluster data}

SIFiltered <- subset(SI, idents = c(1,2,3,4,5,6))

DefaultAssay(SIFiltered) <- "SCT"

SIFiltered <- RunPCA(SIFiltered, verbose = FALSE)
ElbowPlot(SIFiltered)
SIFiltered <- RunUMAP(SIFiltered, reduction = "pca", dims = 1:25)

SIFiltered <- FindNeighbors(SIFiltered, reduction = "pca", dims = 1:25)
SIFiltered <- FindClusters(SIFiltered, resolution = 0.3)

DimPlot(SIFiltered, label =T)

```

#plot marker genes
```{r plot marker genes}

DefaultAssay(SIFiltered) <- "RNA"

FeaturePlot(SIFiltered, c( "Pdgfra", "Myh11",
                     "Foxl1", "Bmp5", "Bmp7", "Lama5", "Adamts18", "Wif1",
                     "Vcam1", "Ccl8", "Wnt4", "Tnc", "Cd55",
                     "Cd34", "Pi16", "Cd81", "Pcolce2", "Rspo2", "Grem1"), order = T)

```


#name clusters
```{r name clusters}

ClusterIDs <- c( "SAF", "bottomMAF", "SEMF", "topMAF", "myocyte")

names(ClusterIDs) <- levels(SIFiltered)
SIFiltered <- RenameIdents(SIFiltered, ClusterIDs)

SIFiltered@active.ident <- factor(SIFiltered@active.ident,
                                       levels = c( "SEMF", "topMAF",  "bottomMAF", "SAF","myocyte"))

DimPlot(SIFiltered, cols = c( "SEMF" = "#F6A937", "topMAF" = "#1DE0FC", "bottomMAF" = "#1D7BFC", 'SAF' = '#129B93', "myocyte" = "#C51B7C")) & NoAxes()

saveRDS(SIFiltered, "/dir/Mm_SI_this_study.rds")

```























