---
title: "Code2_mouse_integration_scRNAseq"
output: html_document
date: "2025-07-14"
---

#load libraries
```{r}

library(Seurat)
library(tidyverse)
library(patchwork)
library(biomaRt)
library(RColorBrewer)

```


#create Seurat objects functions
```{r}
SeuratAddCov <- function(SeuObj, species){
  
  if(species == "mmu"){
    data <- PercentageFeatureSet(SeuObj, pattern="^mt-", col.name = "percent.mt")
    data <- PercentageFeatureSet(data, pattern="^Rp[l|s][[:digit:]]", col.name="percent.rb")  
    data <- PercentageFeatureSet(data, pattern="^Rn45s", col.name="percent.Rn45s")  
    
  }else if(species == "hs"){
    data <- PercentageFeatureSet(SeuObj, pattern="^MT-", col.name = "percent.mt")
    data <- PercentageFeatureSet(data, pattern="^RP[L|S][[:digit:]]", col.name="percent.rb")
  }
  return(data)
}


SeuratCreateObj <- function(dir, name){
  data <- Read10X(dir)
  SeuObj <- CreateSeuratObject(data, name, min.cells = 3, min.features = 200, names.field = 2, names.delim = "_")
  SeuObj <- SeuratAddCov(SeuObj, "mmu")
  saveRDS(SeuObj, paste0(name, ".rds"))
  return(SeuObj)
}

```

#creaate objects
```{r}
#Roulis

C1A <- read.table(file = "/dir/GSM4227211_C1A_gene_exon_tagged_dge.txt.gz", sep = "\t", header = T, row.names = 1)
SeuObjC1A <- CreateSeuratObject(counts = C1A, project = "C1A")
SeuObjC1A[["percent.mt"]] <- PercentageFeatureSet(SeuObjC1A, pattern = "^mt-")
SeuObjC1A[["percent.rb"]] <- PercentageFeatureSet(SeuObjC1A, pattern = "^Rp[l|s][[:digit:]]")
saveRDS(SeuObjC1A, "RoulisC1A.rds")


C2A <- read.table(file = "/dir/GSM4227211_C1A_gene_exon_tagged_dge.txt.gz", sep = "\t", header = T, row.names = 1)
SeuObjC2A <- CreateSeuratObject(counts = C2A, project = "C1A")
SeuObjC2A[["percent.mt"]] <- PercentageFeatureSet(SeuObjC2A, pattern = "^mt-")
SeuObjC2A[["percent.rb"]] <- PercentageFeatureSet(SeuObjC2A, pattern = "^Rp[l|s][[:digit:]]")
saveRDS(SeuObjC2A, "RoulisC2A.rds")

```

```{r}
#Jasso

if(!file.exists("Jasso.rds")){
  dir <- ("/dir/")
  SeuObjJasso <- SeuratCreateObj(dir, "Jasso")
  
}else{
  SeuObjJasso <- readRDS("Jasso.rds")
}

metadata <- read_tsv("metadata.tsv", col_names = T)
metadata <- metadata %>% remove_rownames %>% column_to_rownames(var="cell_barcode")

SeuObjJasso <- AddMetaData(SeuObjJasso, metadata)
H2O <- subset(SeuObjJasso, Treatment == "H2O")
saveRDS(H2O, "/dir/Jasso.rds")

```

```{r}
#Kinchen

Kinchen <- read.table(file = "/dir/GSE114374_Mouse_HC_expression_matrix.txt.gz", sep = "\t", header = T, row.names = 1)
SeuObjKinchen <- CreateSeuratObject(counts = Kinchen, project = "MmHC", min.cells = 3, min.features = 200)
SeuObjKinchen[["percent.mt"]] <- PercentageFeatureSet(SeuObjKinchen, pattern = "^mt-")
SeuObjKinchen[["percent.rb"]] <- PercentageFeatureSet(SeuObjKinchen, pattern = "^Rp[l|s][[:digit:]]")
saveRDS(SeuObjKinchen, "/dir/Kinchen.rds")

```

```{r}
#Brügger

Brügger <- read.table(file = "/dir/GSM4569728_wt_colon_read_counts_raw.txt.gz", sep = "\t", header = T, row.names = 1)
SeuObjBrügger <- CreateSeuratObject(counts = Brügger, project = "Healthy",  min.cells = 3, min.features = 200)
SeuObjBrügger[["percent.mt"]] <- PercentageFeatureSet(SeuObjBrügger, pattern = "^mt-")
SeuObjBrügger[["percent.rb"]] <- PercentageFeatureSet(SeuObjBrügger, pattern = "^Rp[l|s][[:digit:]]")
SeuObjBrügger@meta.data[["orig.ident2"]] <- "Brugger"
saveRDS(SeuObjBrügger, "/dir/Brugger.rds")

```


#filter Seurat objects
```{r}

RoulisC1A <- readRDS("/dir/RoulisC1A.rds")
RoulisC1A@meta.data[["orig.ident"]] <- "Roulis"

VlnPlot(RoulisC1A, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC1A, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

RoulisC1A <- subset(RoulisC1A, subset =
                       nFeature_RNA > 150 & nFeature_RNA < 1500 & 
                       nCount_RNA > 150 & nCount_RNA < 3000 & 
                       percent.mt < 15 & percent.rb < 25)

VlnPlot(RoulisC1A, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC1A, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

RoulisC2A <- readRDS("/dir/RoulisC2A.rds")
RoulisC2A@meta.data[["orig.ident"]] <- "Roulis"


VlnPlot(RoulisC2A, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC2A, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

RoulisC2A <- subset(RoulisC2A, subset =
                       nFeature_RNA > 150 & nFeature_RNA < 1500 & 
                       nCount_RNA > 150 & nCount_RNA < 3000 & 
                       percent.mt < 15 & percent.rb < 25)

VlnPlot(RoulisC2A, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(RoulisC2A, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

Jasso <- readRDS("/dir/Jasso.rds")
Jasso@meta.data[["orig.ident"]] <- "Jasso"

VlnPlot(Jasso, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Jasso, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

Jasso <- subset(Jasso, subset =
                       nFeature_RNA > 600 & nFeature_RNA < 2500 & 
                       nCount_RNA > 1000 & nCount_RNA < 6000 & 
                       percent.mt < 5 & percent.rb < 25)

VlnPlot(Jasso, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Jasso, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

Kinchen <- readRDS("/dir/Kinchen.rds")
Kinchen@meta.data[["orig.ident"]] <- "Kinchen"

VlnPlot(Kinchen, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Kinchen, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")


Kinchen <- subset(Kinchen, subset =
                       nFeature_RNA > 500 & nFeature_RNA < 3500 & 
                       nCount_RNA > 2000 & nCount_RNA < 3500 & 
                       percent.mt < 2 & percent.rb < 12)

VlnPlot(Kinchen, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Kinchen, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

Brugger <- readRDS("/dir/Brugger.rds")
Brugger@meta.data[["orig.ident"]] <- "Brügger"

VlnPlot(Brugger, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Brugger, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

Brugger <- subset(Brugger, subset =
                       nFeature_RNA > 1500 & nFeature_RNA < 4000 & 
                       nCount_RNA > 7500 & nCount_RNA < 10500 & 
                       percent.mt < 20 & percent.rb < 30)

VlnPlot(Brugger, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(Brugger, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

```{r}

ThisStudy <- readRDS("/dir/SeuObjColon.rds")
ThisStudy@meta.data[["orig.ident"]] <- "ThisStudy"

VlnPlot(ThisStudy, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(ThisStudy, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

ThisStudy <- subset(ThisStudy, subset = nFeature_RNA > 1000 & nFeature_RNA < 4000
                         & nCount_RNA < 13000 & nCount_RNA > 1000
                         & percent.mt < 6
                         & percent.rb < 25 & percent.rb > 3)

VlnPlot(ThisStudy, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)

VlnPlot(ThisStudy, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
FeatureScatter(ThisStudy, feature1 ="nFeature_RNA", feature2 =  "nCount_RNA")

```

#normalize and integrate data, run PCA
```{r}

datalist <- list(RoulisC1A2, RoulisC2A2 , JassoH2O, KinchenHC, Brugger, Koliaraki)

convertHumanGeneList <- function(x){
  require("biomaRt")
  human = useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://dec2021.archive.ensembl.org/")
  mouse = useEnsembl("ensembl", dataset = "mmusculus_gene_ensembl", host = "https://dec2021.archive.ensembl.org/")
  genes = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = x , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)
  humanx <- unique(genes[, 2])
  print(head(humanx))
  return(humanx)
}

m.s.genes <- convertHumanGeneList(cc.genes$s.genes)
m.g2m.genes <- convertHumanGeneList(cc.genes$g2m.genes)

for (i in 1:length(datalist)) {
    datalist[[i]] <- SCTransform(datalist[[i]])
    datalist[[i]] <- CellCycleScoring(datalist[[i]], g2m.features = m.g2m.genes, s.features= m.s.genes)
    datalist[[i]] <- SCTransform(datalist[[i]], vars.to.regress = c( "percent.mt", "percent.rb", "S.Score", "G2M.Score"))
    }

features <- SelectIntegrationFeatures(object.list = datalist, nfeatures = 3000)

datalist <- PrepSCTIntegration(object.list = datalist, anchor.features = features)

datalist <- lapply(X = datalist, FUN = RunPCA, features = features)

anchors <- FindIntegrationAnchors(object.list = datalist, normalization.method = "SCT",
                                  anchor.features = features, reduction = "rpca", k.anchor = 7)

DataIntegrated <- IntegrateData(anchorset = anchors, normalization.method = "SCT")



DefaultAssay(DataIntegrated) <- "integrated"

DataIntegrated <- RunPCA(DataIntegrated, verbose = FALSE)
ElbowPlot(DataIntegrated)
DataIntegrated <- RunUMAP(DataIntegrated, reduction = "pca", dims = 1:30)

DataIntegrated <- FindNeighbors(DataIntegrated, reduction = "pca", dims = 1:30)
DataIntegrated <- FindClusters(DataIntegrated, resolution = 0.7)

DefaultAssay(DataIntegrated) <- "RNA"
DataIntegrated <- NormalizeData(DataIntegrated, verbose = F)
DataIntegrated <- FindVariableFeatures(DataIntegrated, selection.method = "vst", nfeatures = 3000, verbose = FALSE)
DataIntegrated <- ScaleData(DataIntegrated, verbose = FALSE)

DimPlot(DataIntegrated, reduction = "umap", label = T, label.size = 4)

```

#split data by original dataset
```{r}
DimPlot(DataIntegrated, reduction = "umap", label = F, split.by = "orig.ident")
```

#plot cell type markers
```{r}

DefaultAssay(DataIntegrated) <- "RNA"

FeaturePlot(DataIntegrated, c( "Pdgfra", "Acta2", "Epcam", "Lgr5", "Muc2",  "Ptprc", "Pecam1", "Lyve1", "Notch3", "Ano1", "Kit", "S100b", "Tubb3" ), order = T)

```

#check cluster quality
```{r}

FeaturePlot(DataIntegrated, c("nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rb"), order =T)

```

#filter out immune and poor quality cells
```{r}

DataIntegratedSub <- subset(DataIntegrated, idents = c(0,1,2,3,4,5,6,7,8,9,10,12,13,14,15,17))

DefaultAssay(DataIntegratedSub) <- "integrated"

DataIntegratedSub <- RunPCA(DataIntegratedSub, verbose = FALSE)
DataIntegratedSub <- RunUMAP(DataIntegratedSub, reduction = "pca", dims = 1:15)

DataIntegratedSub <- FindNeighbors(DataIntegratedSub, reduction = "pca", dims = 1:15)
DataIntegratedSub <- FindClusters(DataIntegratedSub, resolution = 0.6)


DimPlot(DataIntegratedSub, reduction = "umap", label = T, pt.size = 0.1)

```

#find marker genes
```{r}

Markers <- FindAllMarkers(DataIntegratedSub, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.25, assay = "RNA")

top10 <- Markers %>%
  group_by(cluster)%>%
  top_n(n = 10, wt = avg_log2FC)

DefaultAssay(DataIntegratedSub) <- "RNA"
VlnPlot(DataIntegratedSub, features =  top10$gene, pt.size = 0)

```

#name clusters
```{r}

ClusterIDs <- c("Dist_basal_MAF", "SAF", "Basal_SEMF", "Prox_Basal_MAF", "SMC", "Luminal_SEMF", "VEC", "Colonocytes", "Luminal_MAF", "LEC","Goblet", "Glia",  "Peri", "MIF", "Peri&EC")

names(ClusterIDs) <- levels(DataIntegratedSub)
DataIntegrated2 <- RenameIdents(DataIntegratedSub, ClusterIDs)

DataIntegrated2@active.ident <- factor(DataIntegrated2@active.ident,
                                       levels = c( "Luminal_SEMF", "Basal_SEMF", "Luminal_MAFF",  "Prox_Basal_MAF", "Dist_basal_MAF", "SAF", "MIF",
                                                   "SMC", "Peri","Peri&EC", "VEC", "LEC", "Glia", "Colonocytes", "Goblet"))

DimPlot(DataIntegrated2, reduction = "umap", label = F, pt.size = 0.5,
       cols = c("Luminal_SEMF" = "#FCED1D", "Basal_SEMF" = "#F6A937", "Luminal_MAF" = "#1DE0FC", "Prox_basal_MAF" = "#1D7BFC", "Dist_basal_MAF" = "#182FAF", 'SAF' = '#129B93',  "MIF" = "#035C56", 
                "SMC" = "#C51B7C","Peri" = "#6E5F57", "Peri&EC" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E", "Glia" = "#C2D9F0", "Colonocytes" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#split data by original dataset
```{r}

DimPlot(DataIntegrated2, reduction = "umap", label = F, pt.size = 0.1, split.by = "orig.ident",
       cols = c("Luminal_SEMF" = "#FCED1D", "Basal_SEMF" = "#F6A937", "Luminal_MAF" = "#1DE0FC", "Prox_basal_MAF" = "#1D7BFC", "Dist_basal_MAF" = "#182FAF", 'SAF' = '#129B93',  "MIF" = "#035C56", 
                "SMC" = "#C51B7C","Peri" = "#6E5F57", "Peri&EC" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E", "Glia" = "#C2D9F0", "Colonocytes" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#split data by sampling site
```{r}

DimPlot(DataIntegrated2, split.by = "Site", pt.size = 0.2,
      cols = c("Luminal_SEMF" = "#FCED1D", "Basal_SEMF" = "#F6A937", "Luminal_MAF" = "#1DE0FC", "Prox_basal_MAF" = "#1D7BFC", "Dist_basal_MAF" = "#182FAF", 'SAF' = '#129B93',  "MIF" = "#035C56", 
                "SMC" = "#C51B7C","Peri" = "#6E5F57", "Peri&EC" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E", "Glia" = "#C2D9F0", "Colonocytes" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#find markers between fibroblast populations
```{r}

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Luminal_SEMF",
                       ident.2 = c("Basal_SEMF", "Luminal_MAF", "Prox_basal_MAF", "Dist_basal_MAF", "SAF", "MIF", "SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Luminal_SEMF",
                       ident.2 = c("Basal_SEMF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")


Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Basal_SEMF",
                       ident.2 = c("Luminal_SEMF", "Luminal_MAF", "Prox_basal_MAF", "Dist_basal_MAF", "SAF", "MIF", "SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Basal_SEMF",
                       ident.2 = c("Luminal_SEMF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")




Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Luminal_SEMF",
                       ident.2 = c("top_SEMF", "Basal_SEMF", "Prox_basal_MAF", "Dist_basal_MAF", "SAF", "MIF", "SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.35, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Luminal_SEMF",
                       ident.2 = c( "Prox_basal_MAF", "Dist_basal_MAF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.35, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers1 <- FindMarkers(DataIntegrated2,
                       ident.1 = "prox_bottom_mucF",
                       ident.2 = c("dist_bottom_mucF"),
                       logfc.threshold = 0.3, min.diff.pct = 0.3, min.pct = 0.3, only.pos = T, assay ="RNA")


Markers2 <- FindMarkers(DataIntegrated2,
                       ident.1 = "dist_bottom_mucF",
                       ident.2 = c("prox_bottom_mucF"),
                       logfc.threshold = 0.3, min.diff.pct = 0.3, min.pct = 0.3, only.pos = T, assay ="RNA")


Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Dist_basal_MAF",
                       ident.2 = c("Luminal_SEMF", "Luminal_MAF", "Prox_basal_MAF", "Basal_SEMF", "SAF", "MIF", "SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.4, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "Dist_basal_MAF",
                       ident.2 = c( "Luminal_MAF", "Prox_basal_MAF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.3, min.pct = 0.5, only.pos = T, assay ="RNA")




Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "SAF",
                       ident.2 = c("Luminal_SEMF", "Luminal_MAF", "Prox_basal_MAF", "Dist_basal_MAF","Basal_SEMF", "MIF", "SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "MIF",
                       ident.2 = c("Luminal_SEMF", "Luminal_MAF", "Prox_basal_MAF", "Dist_basal_MAF", "SAF", "Basal_SEMF", "SMC"),
                       logfc.threshold = 0.5, min.diff.pct = 0.5, min.pct = 0.5, only.pos = T, assay ="RNA")


Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = "SMC",
                       ident.2 = c("Luminal_SEMF", "Luminal_MAF", "Prox_basal_MAF", "Dist_basal_MAF", "SAF", "Basal_SEMF", "MIF"),
                       logfc.threshold = 0.5, min.diff.pct = 0.55, min.pct = 0.5, only.pos = T, assay ="RNA")


VlnPlot(DataIntegrated2, rownames(Markers), pt.size = 0)

```

#plot validated and new fibroblast subpopulation markers
```{r}

DataIntegrated2@active.ident <- factor(DataIntegrated2@active.ident,
                                       levels = c("Goblet","Colonocytes", "Glia","LEC","VEC", "Peri&EC","Peri", "SMC",
                                                  "MIF", "SAF", "Dist_basal_MAF", "Prox_basal_MAF","Luminal_MAF",  "Basal_SEMF", "Luminal_SEMF"))

genes <- c("Pdgfra", "Acta2", "Des", "Foxl1", "F3", "Procr", "Bmp5", "Bmp7" , "Wnt5a", "Wnt4", "Fgfr2", "Igfbp5", "Col15a1", "Cd34", "Cd81", "Wnt2b", "Grem1", "Grem2", "Rspo3", "Lama5", "Wif1", "Pi16", "Kcnn3")

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size", col.min = 0)+
  theme(axis.text=element_text(angle=90))

```

#plot markers for non-fibroblast cell types
```{r}

genes <- c( "Acta2",  "Lmod1",  "Cnn1", "Tagln", "Myh11", "Myl9",  "Mylk",  "Actg2","Des",
            "Pdgfrb", "Cspg4" , "Rgs4", "Vtn", "Notch3",
            "Flt1","Pecam1", 
            "Lyve1",  "Flt4", "Prox1",
            "S100b", "Gfap", "Uchl1",
            "Epcam", "Aldob", "Fabp2", "Muc2")

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,  genes, scale.by = "size")+
    theme(axis.text=element_text(angle=90))

```

#plot signatures for stem cell niche factors
```{r}

genelist1 <- list(c("Grem1", "Grem2", "Rspo3"))

DataIntegrated2 <- AddModuleScore(object = DataIntegrated2, features = genelist1, name = "genelist1")

FeaturePlot(object = DataIntegrated2, features = "genelist1", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#plot signatures for differentiation niche factors
```{r}

genelist2 <- list(c("Bmp5", "Bmp7", "Wnt5a"))

DataIntegrated2 <- AddModuleScore(object = DataIntegrated2, features = genelist2, name = "genelist2")

FeaturePlot(object = DataIntegrated2, features = "genelist2", order = T) & scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral"))) & NoAxes()

```

#find shared and species-specific markers for mouse fibroblast subpopulations
```{r}

##SEMF

Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = c("Basal_SEMF", "Luminal_SEMF"),
                       ident.2 = c("Glia","LEC","VEC", "Peri&EC","Peri", "SMC", "MIF", "SAF", "Dist_basal_MAF", "Prox_basal_MAF","Luminal_MAF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")


top50Mm <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#top50SEMF = human SEMF markers
top50Mm$gene <- toupper(rownames(top50Mm))
top50SEMF$gene <- toupper(rownames(top50SEMF))

common_genes <- intersect(top50Mm$gene, top50SEMF$gene)
common_genes


to_mouse_case <- function(x) {
  paste0(substr(x, 1, 1), tolower(substr(x, 2, nchar(x))))
}

common_mouse_format <- to_mouse_case(common_genes)
common_mouse_format


mouse_specific <- top50Mm[!top50Mm$gene %in% common_genes, ]

top5Mm <- mouse_specific %>%
  top_n(n = 5, wt = avg_log2FC)
top5Mm <- rownames(top5Mm)
MmSEMF <- c(common_mouse_format, top5Mm)


##MAF
Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = c( "Dist_basal_MAF", "Prox_basal_MAF","Luminal_MAF"),
                       ident.2 = c("Glia","LEC","VEC", "Peri&EC","Peri", "SMC", "MIF", "SAF", "Basal_SEMF", "Luminal_SEMF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")

top50Mm <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#top50MAF = human MAF markers
top50Mm$gene <- toupper(rownames(top50Mm))
top50MAF$gene <- toupper(rownames(top50MAF))

common_genes <- intersect(top50Mm$gene, top50MAF$gene)
common_genes


to_mouse_case <- function(x) {
  paste0(substr(x, 1, 1), tolower(substr(x, 2, nchar(x))))
}

common_mouse_format <- to_mouse_case(common_genes)
common_mouse_format


mouse_specific <- top50Mm[!top50Mm$gene %in% common_genes, ]

top5Mm <- mouse_specific %>%
  top_n(n = 5, wt = avg_log2FC)
top5Mm <- rownames(top5Mm)
MmMAF <- c(common_mouse_format, top5Mm)


##SAF
Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = c("SAF"),
                       ident.2 = c("Glia","LEC","VEC", "Peri&EC","Peri", "SMC", "MIF", "Basal_SEMF", "Luminal_SEMF", "Dist_basal_MAF", "Prox_basal_MAF","Luminal_MAF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")


top50Mm <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#top50SAF = human SAF markers
top50Mm$gene <- toupper(rownames(top50Mm))
top50SAF$gene <- toupper(rownames(top50SAF))

common_genes <- intersect(top50Mm$gene, top50SAF$gene)
common_genes


to_mouse_case <- function(x) {
  paste0(substr(x, 1, 1), tolower(substr(x, 2, nchar(x))))
}

common_mouse_format <- to_mouse_case(common_genes)
common_mouse_format


mouse_specific <- top50Mm[!top50Mm$gene %in% common_genes, ]

top5Mm <- mouse_specific %>%
  top_n(n = 5, wt = avg_log2FC)
top5Mm <- rownames(top5Mm)
MmSAF <- c(common_mouse_format,top5Mm)



##MIF
Markers <- FindMarkers(DataIntegrated2,
                       ident.1 = c("MIF"),
                       ident.2 = c("Glia","LEC","VEC", "Peri&EC","Peri", "SMC", "SAF", "Basal_SEMF", "Luminal_SEMF", "Dist_basal_MAF", "Prox_basal_MAF","Luminal_MAF"),
                       logfc.threshold = 0.25, min.diff.pct = 0.25, min.pct = 0.25, only.pos = T, assay ="RNA")


top50Mm <- Markers %>%
  top_n(n = 50, wt = avg_log2FC)

#top50MIF = human MIF markers
top50Mm$gene <- toupper(rownames(top50Mm))
top50MIF$gene <- toupper(rownames(top50MIF))

common_genes <- intersect(top50Mm$gene, top50MIF$gene)
common_genes


to_mouse_case <- function(x) {
  paste0(substr(x, 1, 1), tolower(substr(x, 2, nchar(x))))
}

common_mouse_format <- to_mouse_case(common_genes)
common_mouse_format


mouse_specific <- top50Mm[!top50Mm$gene %in% common_genes, ]

top5Mm <- mouse_specific %>%
  top_n(n = 5, wt = avg_log2FC)
top5Mm <- rownames(top5Mm)
MmMIF <- c(common_mouse_format,top5Mm)



MmPlot <- c(MmSEMF, MmMAF, MmSAF, MmMIF)

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,  MmPlot, scale.by = "size", col.min = 0)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#luminal and basal markers
```{r}

VlnPlot(DataIntegrated2, c( "Wnt4", "Grem1"), pt.size = 0,
               cols = c("Luminal_SEMF" = "#FCED1D", "Basal_SEMF" = "#F6A937", "Luminal_MAF" = "#1DE0FC", "Prox_basal_MAF" = "#1D7BFC", "Dist_basal_MAF" = "#182FAF", 'SAF' = '#129B93',  "MIF" = "#035C56", 
                "SMC" = "#C51B7C","Peri" = "#6E5F57", "Peri&EC" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E", "Glia" = "#C2D9F0", "Colonocytes" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#proximal and distal MAF markers
```{r}

VlnPlot(DataIntegrated2, c( "Tnc", "Cd55"), pt.size = 0,
              cols = c("Luminal_SEMF" = "#FCED1D", "Basal_SEMF" = "#F6A937", "Luminal_MAF" = "#1DE0FC", "Prox_basal_MAF" = "#1D7BFC", "Dist_basal_MAF" = "#182FAF", 'SAF' = '#129B93',  "MIF" = "#035C56", 
                "SMC" = "#C51B7C","Peri" = "#6E5F57", "Peri&EC" = "#A89E99", "VEC" = "#E1E0DF", "LEC" = "#47362E", "Glia" = "#C2D9F0", "Colonocytes" = "#C3936D","Goblet" = "#734724")) & NoAxes()

```

#luminal and basal SEMF markers
```{r}

DefaultAssay(DataIntegrated2) <- "RNA"

FeaturePlot(DataIntegrated2, c( "Lama5", "Lgr5", "Adamts18", "Wif1", "Bmp3"), order =T)

```

#plot MIF markers
```{r}

genes <- c("Ramp1", "Kcnn3", "Chodl", "Smoc2", "P2ry1", "Thbs2")

DotPlot(DataIntegrated2, genes, assay = "RNA", dot.min = 0.05, scale.by = "size", col.min = 0)+
    theme(axis.text=element_text(angle=90))

```

#make MIF genelist for GSEA
```{r}

Markers <- FindMarkers(DataIntegrated2, ident.1 = "MIF", min.pct = 0.15, logfc.threshold = 0.1, assay = "RNA")
Markers$gene <- rownames(Markers)
Markers <- Markers %>%
  arrange(desc(avg_log2FC))
writexl::write_xlsx(Markers, "/dir/MIF_markersMF.xlsx")

```

#GSEA GOMF on MIF
```{r}

library(readr)
library(patchwork)
library(clusterProfiler)
library(org.Mm.eg.db)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(cowplot)
library(readxl)
library(tidyverse)


Data <- read_xlsx("/dir/MIF_markersMF.xlsx")
Data <- Data %>%
  dplyr::select(avg_log2FC, gene)
  
genelist <- Data$avg_log2FC
names(genelist) <- Data$gene

genelist <- sort(genelist, decreasing = TRUE)

names(genelist)[duplicated(names(genelist))]

head(genelist, 5)
any(genelist == 0)
any(is.na(genelist))
any(names(genelist) == 0)
any(is.na(names(genelist)))

table(is.finite(genelist)) 

GOMF_set <- msigdbr(species = "Mus musculus", category = "C5", subcategory = "MF")

GOMFTerm2 <- GOMF_set %>% dplyr::select(gs_name, gene_symbol)

GOMF <- GSEA(geneList = genelist,
                              TERM2GENE = GOMFTerm2,
                              minGSSize = 15,
                              pvalueCutoff = 0.05,
                              verbose = TRUE)


p1 <- gseaplot2(GOMF, geneSetID = "GOMF_VOLTAGE_GATED_MONOATOMIC_CATION_CHANNEL_ACTIVITY", title = "GOMF_VOLTAGE_GATED_MONOATOMIC_CATION_CHANNEL_ACTIVITY")
p2 <- gseaplot2(GOMF, geneSetID = "GOMF_POTASSIUM_CHANNEL_ACTIVITY", title = "GOMF_POTASSIUM_CHANNEL_ACTIVITY")
p3 <- gseaplot2(GOMF, geneSetID = "GOMF_VOLTAGE_GATED_CHANNEL_ACTIVITY", title = "GOMF_VOLTAGE_GATED_CHANNEL_ACTIVITY")
p4 <- gseaplot2(GOMF, geneSetID = "GOMF_MONOATOMIC_CATION_CHANNEL_ACTIVITY", title = "GOMF_MONOATOMIC_CATION_CHANNEL_ACTIVITY")
p5 <- gseaplot2(GOMF, geneSetID = "GOMF_G_PROTEIN_COUPLED_RECEPTOR_ACTIVITY", title = "GOMF_G_PROTEIN_COUPLED_RECEPTOR_ACTIVITY")
p6 <- gseaplot2(GOMF, geneSetID = "GOMF_METAL_ION_TRANSMEMBRANE_TRANSPORTER_ACTIVITY", title = "GOMF_METAL_ION_TRANSMEMBRANE_TRANSPORTER_ACTIVITY")

gg_p1 <- grid::grid.grabExpr({print(p1)})
gg_p2 <- grid::grid.grabExpr({print(p2)})
gg_p3 <- grid::grid.grabExpr({print(p3)})
gg_p4 <- grid::grid.grabExpr({print(p4)})
gg_p5 <- grid::grid.grabExpr({print(p5)})
gg_p6 <- grid::grid.grabExpr({print(p6)})

wrap_plots(list(gg_p1, gg_p2, gg_p3, gg_p4, gg_p5, gg_p6))

```

#subset and recluster SMCs
```{r}

DataIntegrated2Sub <- subset(DataIntegrated2, idents = c("SMC"))

DefaultAssay(DataIntegrated2Sub) <- "integrated"

DataIntegrated2Sub <- RunPCA(DataIntegrated2Sub, verbose = FALSE)
DataIntegrated2Sub <- RunUMAP(DataIntegrated2Sub, reduction = "pca", dims = 1:10)

DataIntegrated2Sub <- FindNeighbors(DataIntegrated2Sub, reduction = "pca", dims = 1:10)
DataIntegrated2Sub <- FindClusters(DataIntegrated2Sub, resolution = 0.075)

DimPlot(DataIntegrated2Sub, reduction = "umap", label = T)

```

#find SMC markers
```{r}

DefaultAssay(DataIntegrated2Sub) <- "RNA"

Markers <- FindAllMarkers(DataIntegrated2Sub, only.pos = TRUE, min.pct = 0.4, logfc.threshold = 0.4, min.diff.pct = 0.4, assay = "RNA")

FeaturePlot(DataIntegrated2, rownames(Markers), order =T) & NoAxes()

```

#rename SMC clusters
```{r}

ClusterIDs <- c( "myocytes", "MM_SMC")
names(ClusterIDs) <- levels(DataIntegrated2Sub)
DataIntegrated2Sub <- RenameIdents(DataIntegrated2Sub, ClusterIDs)

DimPlot(DataIntegrated2Sub, reduction = "umap", label = F, cols = c("myocytes" = "#D54B76","MM_SMC" = "#F6AEE5"), pt.size = 0.7) & NoAxes()

```

#modify function for stacked vlnplot
```{r}

modify_vlnplot<- function(obj, 
                          feature, 
                          pt.size = 0, 
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  p<- VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    xlab("") + ylab(feature) + ggtitle("") + 
    theme(legend.position = "none", 
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(), 
          axis.title.y = element_text(size = rel(1), angle = 0), 
          axis.text.y = element_text(size = rel(1)), 
          plot.margin = plot.margin ) 
  return(p)
}

extract_max<- function(p){
  ymax<- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}

StackedVlnPlot<- function(obj, features,
                          pt.size = 0, 
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  plot_list<- purrr::map(features, function(x) modify_vlnplot(obj = obj,feature = x, ...))
  plot_list[[length(plot_list)]]<- plot_list[[length(plot_list)]] +
  theme(axis.text.x=element_text(), axis.ticks.x = element_line())
  ymaxs<- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x + 
                            scale_y_continuous(breaks = c(y)) + 
                            expand_limits(y = y))
  p<- patchwork::wrap_plots(plotlist = plot_list, ncol = 1)
  return(p)
}

```

#plot SMC markers from literature
```{r}

DataIntegrated2Sub@active.ident <- factor(DataIntegrated2Sub@active.ident, levels = c( "MM_SMC","myocytes"))

genes <- c("Acta2", "Myh11", "Actg2", "Des", "Hhip", "Kit")

StackedVlnPlot(DataIntegrated2Sub, genes,  cols = c("myocytes" = "#D54B76","MM SMC" = "#F6AEE5"))

```

#find SMC subtype markers and plot in whole integrated data to find specific ones
```{r}

Markers <- FindAllMarkers(DataIntegrated2Sub, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, min.diff.pct = 0.25, assay = "RNA")

top20Mm <- Markers %>%
  group_by(cluster) %>%
  top_n(n = 20, wt = avg_log2FC)

DotPlot(DataIntegrated2, assay = "RNA", dot.min = 0.05,   top10Mm$gene, scale.by = "size", col.min = 0)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#plot SMC subtype markers in subsetted data
```{r, fig.height=4, fig.width=15}

DotPlot(DataIntegrated2Sub, assay = "RNA", dot.min = 0.05,
        c("Hhip", "Rbp4", "Cpxm2", "Casq2", "Gas1", "Dnaic1", 
          "Grem2", "Rgs5", "Rspo3", "Cxcl12", "Ccl11", "Ifitm1"), scale.by = "size", col.min = 0)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

